---
title: "NOAA Restoration Report"
author: "Ian Combs"
date: "2024-01-18"
output: html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
knitr::opts_knit$set(root.dir = "../data")
options(width = 88)
library(magrittr)
```





### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

<!-- this is where the DOI would go  [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3675991.svg)](https://doi.org/10.5281/zenodo.3675991)
-->


#### [GitHub repository](https://github.com/icombs2017/photogrammetryNOAA){target="_blank"}
###

***
This is the working analysis pipeline to analyze data generated from outplant monitoring for the NFWF EDR grant

***

### All analyses performed with R version `r getRversion()`


# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 


```{r,packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "googlesheets4", "dplyr", "officer","reshape2", "stringr", "flextable", "gridExtra", "ggpubr", "Rmisc", "rcompanion", "RColorBrewer", "googledrive", "gdata", "readxl", "DescTools","patchwork", "FSA", "rstatix", "tidyverse", "lme4", 'PMCRMplus', "EnvStats", "emmeans", "MuMIn", "sjstats", "lmerTest", "gargle", "FSA", "vegan", "gtools", "lubridate")

```


# Importing Data
***
## We are downloading this dataset from our GoogleDrive folder. We will be using the package `googledrive`. Each GoogleDrive file has a unique ID that does not change throughout the lifespan of the document, even if the file name is changed. This ID is housed in the file's URL. 
Example: docs.google.com/spreadsheets/d/FILE_ID_GOES_HERE/other_information/. Below you will copy and paste that into the function `drive_download` within `as_id`. This will save the file locally in the specified path (in this case our data folder) and you will import the folder as you normally would. Downloading it this way decreases the potential human error when downloading, moving folders, renaming, saving etc and ensures that the most up to date file is being utilized. 

# Here we are importing data for our plug counts from google drive

```{r, plugLoadingData, include = TRUE}
# 
#  data <- drive_download(
#  as_id("1b4IWK7tfSt3n8o8NlgJKSDoqj_QkHvKcaJBNz85RBv4"),
#    path = "../data/outplantData.xlsx",
#    overwrite = TRUE)



# APAL Data
apalData <- read_excel("../data/outplantData.xlsx", sheet = "ex-situ Apal outplants", na = c("", "N/A", "NA"))

# ACER Data
acerData <- read_excel("../data/outplantdata.xlsx", sheet = "in-situ Acer outplants", na = c("", "N/A", "NA"))


# Ex-Situ Massives Data, theres some weird data artifacts in the beginnnig of the data set so skipping a few columns just to get to the data we need
exMsvsDataNames <- names(read_excel("../data/outplantdata.xlsx", sheet = "ex-situ Massive outplants", na = c("", "N/A", "NA"), col_names = TRUE))
exMsvsData <- read_excel("../data/outplantdata.xlsx", sheet = "ex-situ Massive outplants", na = c("", "N/A", "NA", "N/P", "#N/A", "#REF!"),skip = 28, col_names = exMsvsDataNames)

# In-Situ Massives Data
inMsvsData <- read_excel("../data/outplantdata.xlsx", sheet = "in-situ Massive outplants", na = c("", "N/A", "NA"))

```

# APAL Monitoring summary

## Time to do some data wrangling for the APAL dataset. Since we only care about the NOAA right now, we are going to filter the data set to just include those given to the grant *NOAA*. Also including a fixed string called apalDeliverables and apalCumulativeDeliverables that includes what each deliverable is each year and cumulatively by year.


```{r, apalDataWrangling, include = TRUE}


outplantAPAL <- apalData %>% 
  select(Year, 
         Date, 
         `Site ID`, 
         Deliverable, 
         `# plugs`, 
         `1-month Monitoring`,
         `1-month Survival`, 
         `1-month % Dead`, 
         `1-month % Missing`,
         `1-month % Plot Survival`, 
         `% of Plots with Partial Mortality`,
         `Avg. % Partial Mortality within Affected Plots`,
         `12-month Monitoring`,                           
         `12-month Survival`,                             
         `12-month % Dead`,                               
         `12-month % Missing`,                            
         `12-month % Plot Survival`) %>% 
  filter(Deliverable %in% c("NOAA Yr1", "NOAA Yr2", "NOAA Yr3", "NOAA Yr4")) %>% 
  mutate_at(c("Year", 
         "Site ID", 
        "Deliverable"), as.factor) %>% 
  droplevels() %>% 
  mutate(cum_sum = cumsum(`# plugs`)) %>% 
  mutate(Date = as.Date(Date, "%m %d %Y"))


outplantByYearAP <- outplantAPAL %>% 
  select(Year, 
         Date, 
         `Site ID`, 
         Deliverable, 
         `# plugs`) %>% 
  group_by(Deliverable) %>% 
  mutate(totalOutplanted = sum(`# plugs`))

apalDeliverables <- c(2800, 2800, 1400, 1400)




apalCumulativeDeliverables <- c(2800, 5600, 7000, 8400)






```

# Plotting APAL Outplant Numbers

## Plotting outplant numbers for all APAL over the course of the three years.


```{r, apalOutplantCumulativePlot, include = TRUE}

outplantAPALplot1<- ggplot(outplantAPAL, aes(x = Date, y = cum_sum))+
  geom_line(aes(color = "Fragments Outplanted", linetype = 'solid'), linewidth = 3)+
  geom_hline(aes(yintercept = apalCumulativeDeliverables[1], linetype = 'Year 1 Outplant Goal', color = 'Year 1 Outplant Goal'), linewidth = 3) +
  geom_hline(aes(yintercept = apalCumulativeDeliverables[2], linetype = 'Year 2 Outplant Goal', color = 'Year 2 Outplant Goal'), linewidth = 3)+
geom_hline(aes(yintercept = apalCumulativeDeliverables[3], linetype = 'Year 3 Outplant Goal', color = 'Year 3 Outplant Goal'), linewidth = 3) +
  geom_hline(aes(yintercept = apalCumulativeDeliverables[4], linetype = 'Year 4 Outplant Goal', color = 'Year 4 Outplant Goal'), linewidth = 3) +
  geom_segment(aes(x = as.Date("2023-05-20"), y = 7769,
                   xend = as.Date("2024-03-01"), yend = 7769),
               linetype = 'solid', color = 'blue', size = 3)+ # remove this with full dataset
  scale_y_continuous(limits = c(0, 10000), breaks = seq(0, 30000, by = 2000))+
  ggtitle(expression("Number of "*italic("Acropora palmata")*" Fragments Outplanted per Year"))+
  xlab(" ")+
  ylab("Number of fragments outplanted")+
  scale_color_manual(values = c('blue', 'green3', "gold3", 'purple3', "red3"), name = " ") +
  scale_linetype_manual(name = " ", values = c('solid', 'dashed', 'dashed', 'dashed', 'dashed'), guide = "none")+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "1 month")




outplantAPALplot <- outplantAPALplot1 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))


ggsave("../figures/cumulativeAPALoutplantPlot.png", plot =outplantAPALplot, width = 20, height = 15, units = 'in', dpi = 600)

```
## APAL Outplant Table
NOAA Reports like to have all the data in one table, outplanting, 1 month survival and 12 month survival, so creating a table for all data. 


```{r, apalOutplantTable, include = TRUE}

# Wrangling for the table
outplantAPALtab1 <- apalData %>%
   select(Date, Location, `Site ID`, `# plugs`, Lat, Long, Deliverable,`1-month Monitoring`, `1-month Survival`, `12-month Survival`) %>%
   mutate(`1-month Monitoring` = as.Date(`1-month Monitoring`, `1-month Survival` = as.numeric(`1-month Survival`, `12-month Survival` = as.numeric(`12-month Survival`)))) %>% 
   mutate(Date = as.Date(Date)) %>% 
   filter(Deliverable %in% c("NOAA Yr1", "NOAA Yr2", "NOAA Yr3", "NOAA Yr4")) %>% 
   # select(Date, `Site ID`, `# plugs`, Lat, Long, numberOfGenotypes)
  mutate(oneMonthNumberSurvived = `# plugs` * `1-month Survival`, twelveMonthNumberSurvived = `# plugs` * `12-month Survival`) %>% 
  mutate_at(vars(`1-month Survival`, `12-month Survival`), ~ round(. * 100, 2)) 

#creating templates for "weighted survival"
# 1 month APAL frags
oneMonthWeightedAverageSurvival<- (sum(outplantAPALtab1$oneMonthNumberSurvived))/(sum(outplantAPALtab1$`# plugs`))*100

# 12 month APAL plugs
twelveMonthWeightedAverageSurvival<- (sum(na.omit(outplantAPALtab1$twelveMonthNumberSurvived)))/(sum(na.omit(outplantAPALtab1$`# plugs`)))*100


# Making the table
outplantAPALtab <- outplantAPALtab1 %>%
  select(Date, Location, `Site ID`, `# plugs`, Lat, Long, `1-month Survival`, `12-month Survival`) %>%
    mutate(`12-month Survival` = ifelse(is.na(`12-month Survival`), "NA", `12-month Survival`))%>% 
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Date", value = as_paragraph("Outplant Date")) %>%
  flextable::compose(part = "header", j = "# plugs", value = as_paragraph("Number of Corals")) %>%
  flextable::compose(part = "header", j = "Lat", value = as_paragraph("Latitude")) %>% 
  flextable::compose(part = "header", j = "Long", value = as_paragraph("Longitude")) %>%
  flextable::compose(part = "header", j = "Location", value = as_paragraph("General Location")) %>%
  flextable::compose(part = "header", j = "1-month Survival", value = as_paragraph("1-month % Survival")) %>% 
  flextable::compose(part = "header", j = "12-month Survival", value = as_paragraph("12-month % Survival")) %>% 
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "all") %>% 
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  
    add_footer_row(values = 
    list(Date = "Totals:",
         "", 
         "", 
         `# plugs` = sum(outplantAPALtab1$`# plugs`), 
         "", 
         Long = "Weighted Averages",
         `1-month Survival`= round (oneMonthWeightedAverageSurvival, 2),
         `12-month Survival` = round(twelveMonthWeightedAverageSurvival,2)
         ),
    colwidths = c(1, 1, 1, 1, 1, 1, 1, 1))
 
   
# totals <- c("Totals:", "", sum(outplantAPALtab1$`# plugs`), sum(outplantAPALtab1$numberOfGenotypes))

outplantAPALDoc = read_docx()
outplantAPALDoc = body_add_flextable(outplantAPALDoc, value = outplantAPALtab)
print(outplantAPALDoc, target = "../tables/outplantAPALtable.docx")
outplantAPALtab








```




# ACER Outplant Monitoring
Including the same data wrangling as above but also including a fixed string called acerDeliverables and acerCumulativeDeliverables that includes what each deliverable is each year and cumulatively by year

```{r, acerDataWrangling, include = TRUE}


outplantACER <- acerData %>% 
  select(Year, 
         Date, 
         `Site ID`, 
         Deliverable, 
         `# frags`, 
         `1-month Monitoring`,
         `1-month Survival`, 
         `1-month % Dead`, 
         `1-month % Missing`,
         `1-month % Plot Survival`, 
         `% of Plots with Partial Mortality`,
         `Avg. % Partial Mortality within Affected Plots`,
         `12-month Monitoring`,                           
         `12-month Survival`,                             
         `12-month % Dead`,                               
         `12-month % Missing`,                            
         `12-month % Plot Survival`,
         `% TLE Present...27`) %>% 
  filter(Deliverable %in% c("NOAA Yr1", "NOAA Yr2", "NOAA Yr3", "NOAA Yr4")) %>% 
  mutate_at(c("Year", 
         "Site ID", 
        "Deliverable"), as.factor) %>%
  mutate_at(c("Date"), as.Date) %>% 
  droplevels() %>% 
  # group_by(Deliverable) %>%
  mutate(cum_sum = cumsum(`# frags`)) %>% 
  mutate(Date = as.Date(Date, "%m %d %Y"))
# outplantACER$Date <- as.Date(outplantACER$Date, "%m %d %Y")

outplantByYearAC <- outplantACER %>% 
  select(Year, 
         Date, 
         `Site ID`, 
         Deliverable, 
         `# frags`) %>% 
  group_by(Deliverable) %>% 
  mutate(totalOutplanted = sum(`# frags`))

acerDeliverables <- c(6000, 6000, 3000, 3000)


acerCumulativeDeliverables <- c(6000, 12000, 15000, 18000)


```


## Plotting total number of frags per year with the red dashed line indicating each year's deliverable (every year it is 3285)
```{r, acerOutplantByYearPlot, include = TRUE}

# number of frags per year
outplantACERplot1<- ggplot(outplantByYearAC, aes(x = Deliverable, y = totalOutplanted, fill = Deliverable, color = I('black')))+
  geom_bar(position = "dodge", stat = 'identity', width = 1)+
  geom_hline(yintercept = c(3285,3285,3285), linetype = 'dashed', color =c('red', 'red', 'red'))+
 scale_y_continuous(limits = c(0, 4000), breaks = seq(0, 4000, by = 500))+
  ggtitle("Number of Acropora cervicornis Fragments Outplanted per Year")+
  xlab("Year")+
  ylab("Number of fragments outplanted")

outplantACERplot <- outplantACERplot1 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))




ggsave("../figures/outplantACERplot.png", plot =outplantACERplot, width = 20, height = 15, units = 'in', dpi = 600)


```


# Now plotting cumulative number of ACER Frags per year

```{r, acerOutplantCumulativePlot, include = TRUE}




# cumulative totals
outplantACERplot1<- ggplot(outplantACER, aes(x = Date, y = cum_sum))+
  geom_line(aes(color = "Fragments Outplanted", linetype = 'solid'), linewidth = 3)+
  # geom_point(aes(color = "black"))+
  geom_hline(aes(yintercept = acerCumulativeDeliverables[1], linetype = 'Year 1 Outplant Goal', color = 'Year 1 Outplant Goal'), linewidth = 2) +
  geom_hline(aes(yintercept = acerCumulativeDeliverables[2], linetype = 'Year 2 Outplant Goal', color = 'Year 2 Outplant Goal'), linewidth = 2) +
  geom_hline(aes(yintercept = acerCumulativeDeliverables[3], linetype = 'Year 3 Outplant Goal', color = 'Year 3 Outplant Goal'), linewidth = 2) +
  geom_hline(aes(yintercept = acerCumulativeDeliverables[4], linetype = 'Year 4 Outplant Goal', color = 'Year 4 Outplant Goal'), linewidth = 2) +
  geom_segment(aes(x = as.Date("2023-06-04"), y = 18300,
                   xend = as.Date("2023-12-31"), yend = 18300),
               linetype = 'solid', color = 'blue', size = 3)+ # remove this with full dataset
  scale_y_continuous(limits = c(0, 20000), breaks = seq(0, 20000, by = 1000))+
  ggtitle(expression("Number of "*italic("Acropora cervicornis")*" Fragments Outplanted per Year"))+
  xlab(" ")+
  ylab("Number of fragments outplanted")+
  scale_color_manual(values = c('blue', 'green3', "gold3", 'purple3', "red3"), name = " ") +
  scale_linetype_manual(name = " ", values = c('solid', 'dashed', 'dashed', 'dashed', 'dashed'), guide = "none")+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "1 month")+theme()

 
outplantACERplot <- outplantACERplot1 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))

ggsave("../figures/cumulativeACERoutplantPlot.png", plot =outplantACERplot, width = 20, height = 15, units = 'in', dpi = 600)






```
## ACER Outplant Table


```{r, acerOutplantTable, include = TRUE}
# Wrangling for the table
outplantACERtab1 <- acerData %>%
   select(Date, `General general General Location`, `Site ID`, `# frags`, Lat, Long, Deliverable,`1-month Monitoring`, `1-month Survival`, `12-month Survival`) %>%
   mutate(`1-month Monitoring` = as.Date(`1-month Monitoring`, `1-month Survival` = as.numeric(`1-month Survival`, `12-month Survival` = as.numeric(`12-month Survival`)))) %>% 
   mutate(Date = as.Date(Date)) %>% 
   filter(Deliverable %in% c("NOAA Yr1", "NOAA Yr2", "NOAA Yr3", "NOAA Yr4")) %>% 
   # select(Date, `Site ID`, `# frags`, Lat, Long, numberOfGenotypes)
  mutate(oneMonthNumberSurvived = `# frags` * `1-month Survival`, twelveMonthNumberSurvived = `# frags` * `12-month Survival`) %>% 
  mutate_at(vars(`1-month Survival`, `12-month Survival`), ~ round(. * 100, 2)) 

#creating templates for "weighted survival"
# 1 month ACER frags
oneMonthWeightedAverageSurvival<- (sum(outplantACERtab1$oneMonthNumberSurvived))/(sum(outplantACERtab1$`# frags`))*100

# 12 month ACER frags
twelveMonthWeightedAverageSurvival<- (sum(na.omit(outplantACERtab1$twelveMonthNumberSurvived)))/(sum(na.omit(outplantACERtab1$`# frags`)))*100


# Making the table
outplantACERtab <- outplantACERtab1 %>%
  select(Date, `General Location`, `Site ID`, `# frags`, Lat, Long, `1-month Survival`, `12-month Survival`) %>%
    mutate(`12-month Survival` = ifelse(is.na(`12-month Survival`), "NA", `12-month Survival`))%>% 
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Date", value = as_paragraph("Outplant Date")) %>%
  flextable::compose(part = "header", j = "# frags", value = as_paragraph("Number of Corals")) %>%
  flextable::compose(part = "header", j = "Lat", value = as_paragraph("Latitude")) %>% 
  flextable::compose(part = "header", j = "Long", value = as_paragraph("Longitude")) %>%
  flextable::compose(part = "header", j = "General Location", value = as_paragraph("General Location")) %>%
  flextable::compose(part = "header", j = "1-month Survival", value = as_paragraph("1-month % Survival")) %>% 
  flextable::compose(part = "header", j = "12-month Survival", value = as_paragraph("12-month % Survival")) %>% 
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "all") %>% 
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  
    add_footer_row(values = 
    list(Date = "Totals:",
         "", 
         "", 
         `# frags` = sum(outplantACERtab1$`# frags`), 
         "", 
         Long = "Weighted Averages",
         `1-month Survival`= round (oneMonthWeightedAverageSurvival, 2),
         `12-month Survival` = round(twelveMonthWeightedAverageSurvival,2)
         ),
    colwidths = c(1, 1, 1, 1, 1, 1, 1, 1))
 
   
# totals <- c("Totals:", "", sum(outplantACERtab1$`# frags`), sum(outplantACERtab1$numberOfGenotypes))

outplantACERDoc = read_docx()
outplantACERDoc = body_add_flextable(outplantACERDoc, value = outplantACERtab)
print(outplantACERDoc, target = "../tables/outplantACERtable.docx")
outplantACERtab


```

# Massives Monitoring summary

## Time to do some data wrangling for the Massives dataset. Since we only care about the NOAA right now, we are going to filter the data set to just include those given to the grant *NOAA*. Also including a fixed string called apalDeliverables and apalCumulativeDeliverables that includes what each deliverable is each year and cumulatively by year.Since this includes both ex-situ and in-situ massives we need to comibne those to data sets so we can more easily work with the data. 


```{r, msvsDataWrangling, include = TRUE}

exMsvsData1 <- exMsvsData %>%
  mutate(numberOfGenotypes = rowSums(!is.na(select(., `Genotypes Represented`:ncol(.))))) %>%
  select(1:48, numberOfGenotypes) %>%
  dplyr::rename(totalNumberOfPlugs = `# plugs`) %>%
  filter(Deliverable %in% c("NOAA Yr1", "NOAA Yr2", "NOAA Yr3", "NOAA Yr4")) %>%
 
  mutate_at(c("Site ID",
        "Deliverable"), as.factor) %>%
  mutate_at(c("Date"), as.Date) %>% 
  droplevels() %>% 
  mutate(Date = as.Date(Date, "%m %d %Y")) 

inMsvsData1 <- inMsvsData %>% 
  mutate(numberOfGenotypes = rowSums(!is.na(select(., `Gentoypes Represented at the Site`:ncol(.))))) %>%
  dplyr::rename(totalNumberOfPlugs = `# frags`) %>%
  filter(Deliverable %in% c("NOAA Yr1", "NOAA Yr2", "NOAA Yr3", "NOAA Yr4"))
 


outplantMsvs <- bind_rows(
  mutate(exMsvsData1, source = "exMsvsData1"),
  mutate(inMsvsData1, source = "inMsvsData1")) %>% 
    mutate_at(c("Date"), as.Date) %>% 
  arrange(Date) %>% 
  mutate(cum_sum = cumsum(totalNumberOfPlugs)) %>% 
  pivot_longer(cols = 9:17,
               names_to = "species",
               values_to = "plugsPerSpecies") %>%
  select(Date,
       Location,
       `Site ID`,
       Deliverable,
       source,
       totalNumberOfPlugs,
       species,
       plugsPerSpecies,
       cum_sum,
       numberOfGenotypes,
       `1-month Monitoring`,
       `1-month Survival`,
       `1-month % Dead`,
       `1-month % Missing`,
      `1-month % Plot Survival`,
      `12-month Monitoring`,
       `12-month Survival`,
       `12-month % Dead`,
       `12-month % Missing`,
        `12-month % Plot Survival`) %>%
  mutate_at(c("Site ID",
              "Deliverable", 
              "source",
              "species"), as.factor)

msvsDeliverables <- c(1200, 1200, 600, 600)




msvsCumulativeDeliverables <- c(1200, 2400, 3000, 3600)


```

# Plotting Msvs Outplant Numbers

## Plotting outplant numbers for all Massives over the course of the three years.


```{r, msvsOutplantCumulativePlot, include = TRUE}

outplantMSVSplot1<- ggplot(outplantMsvs, aes(x = Date, y = cum_sum))+
  geom_line(aes(color = "Fragments Outplanted", linetype = 'solid'), linewidth = 3)+
  geom_hline(aes(yintercept = msvsCumulativeDeliverables[1], linetype = 'Year 1 Outplant Goal', color = 'Year 1 Outplant Goal'), linewidth = 3) +
  geom_hline(aes(yintercept = msvsCumulativeDeliverables[2], linetype = 'Year 2 Outplant Goal', color = 'Year 2 Outplant Goal'), linewidth = 3)+
geom_hline(aes(yintercept = msvsCumulativeDeliverables[3], linetype = 'Year 3 Outplant Goal', color = 'Year 3 Outplant Goal'), linewidth = 3) +
  geom_hline(aes(yintercept = msvsCumulativeDeliverables[4], linetype = 'Year 4 Outplant Goal', color = 'Year 4 Outplant Goal'), linewidth = 3) +
  geom_segment(aes(x = as.Date("2023-06-18"), y = 3518,
                   xend = as.Date("2023-12-31"), yend = 3518),
               linetype = 'solid', color = 'blue', size = 3)+ # remove this with full dataset
  scale_y_continuous(limits = c(0, 5000), breaks = seq(0, 5000, by = 500))+
  ggtitle(expression("Number of Massive Coral Fragments Outplanted per Year"))+
  xlab(" ")+
  ylab("Number of fragments outplanted")+
  scale_color_manual(values = c('blue', 'green3', "gold3", 'purple3', "red3"), name = " ") +
  scale_linetype_manual(name = " ", values = c('solid', 'dashed', 'dashed', 'dashed', 'dashed'), guide = "none")+
  scale_x_date(date_labels = "%b-%Y", date_breaks = "1 month")




outplantMSVSplot <- outplantMSVSplot1 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))


ggsave("../figures/cumulativeMSVSoutplantPlot.png", plot =outplantMSVSplot, width = 20, height = 15, units = 'in', dpi = 600)

```
## MSVS Outplant Table
NOAA Reports like to have all the data in one table, outplanting, 1 month survival and 12 month survival, so creating a table for all data. 


```{r, msvsOutplantTable, include = TRUE}

# Wrangling for the table
outplantMSVStab1 <- outplantMsvs %>%
   select(Date, Location, `Site ID`, totalNumberOfPlugs, Deliverable,`1-month Monitoring`, `1-month Survival`, `12-month Survival`) %>%
   mutate(`1-month Monitoring` = as.Date(`1-month Monitoring`, `1-month Survival` = as.numeric(`1-month Survival`, `12-month Survival` = as.numeric(`12-month Survival`)))) %>% 
   mutate(Date = as.Date(Date)) %>% 
   filter(Deliverable %in% c("NOAA Yr1", "NOAA Yr2", "NOAA Yr3", "NOAA Yr4")) %>% 
   # select(Date, `Site ID`, totalNumberOfPlugs, Lat, Long, numberOfGenotypes)
  mutate(oneMonthNumberSurvived = totalNumberOfPlugs * `1-month Survival`, twelveMonthNumberSurvived = totalNumberOfPlugs * `12-month Survival`) %>% 
  mutate_at(vars(`1-month Survival`, `12-month Survival`), ~ round(. * 100, 2)) 

#creating templates for "weighted survival"
# 1 month MSVS frags
oneMonthWeightedAverageSurvival<- (sum(outplantMSVStab1$oneMonthNumberSurvived))/(sum(outplantMSVStab1$totalNumberOfPlugs))*100

# 12 month MSVS plugs
twelveMonthWeightedAverageSurvival<- (sum(na.omit(outplantMSVStab1$twelveMonthNumberSurvived)))/(sum(na.omit(outplantMSVStab1$totalNumberOfPlugs)))*100


# Making the table
outplantMSVStab <- outplantMSVStab1 %>%
  distinct() %>% 
  select(Date, Location, `Site ID`, totalNumberOfPlugs, `1-month Survival`, `12-month Survival`) %>%
    mutate(`12-month Survival` = ifelse(is.na(`12-month Survival`), "NA", `12-month Survival`))%>% 
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Date", value = as_paragraph("Outplant Date")) %>%
  flextable::compose(part = "header", j = "totalNumberOfPlugs", value = as_paragraph("Number of Corals")) %>%
  # flextable::compose(part = "header", j = "Lat", value = as_paragraph("Latitude")) %>% 
  # flextable::compose(part = "header", j = "Long", value = as_paragraph("Longitude")) %>%
  flextable::compose(part = "header", j = "Location", value = as_paragraph("General Location")) %>%
  flextable::compose(part = "header", j = "1-month Survival", value = as_paragraph("1-month % Survival")) %>% 
  flextable::compose(part = "header", j = "12-month Survival", value = as_paragraph("12-month % Survival")) %>% 
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "all") %>% 
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  
    add_footer_row(values = 
    list(Date = "Totals and Weighted Averages:",
         "", 
         "", 
         totalNumberOfPlugs = sum(outplantMSVStab1$totalNumberOfPlugs), 
         `1-month Survival`= round (oneMonthWeightedAverageSurvival, 2),
         `12-month Survival` = round(twelveMonthWeightedAverageSurvival,2)
         ),
    colwidths = c(1, 1, 1, 1, 1, 1))
 
   
# totals <- c("Totals:", "", sum(outplantMSVStab1$totalNumberOfPlugs), sum(outplantMSVStab1$numberOfGenotypes))

outplantMSVSDoc = read_docx()
outplantMSVSDoc = body_add_flextable(outplantMSVSDoc, value = outplantMSVStab)
print(outplantMSVSDoc, target = "../tables/outplantMSVStable.docx")
outplantMSVStab








```















# APAL One Month Monitoring
We're going to wrangle the data so we can plot one month survival




## APAL One Month Data Wrangling
```{r, oneMonthDataWranglingAPAL, include = TRUE}

myColors <- c("red", "yellow", "green")
OneMoSurvivalAPAL2 <- outplantAPAL %>% 
  # mutate_at(c("1-month Survival", "1-month % Dead", "1-month % Missing", "12-month Survival", "12-month % Dead", "12-month % Missing"), as.double) %>%
    pivot_longer(cols = c("1-month Survival", "1-month % Dead", "1-month % Missing"), names_to = "1 month", values_to = "1 mo percentage") %>%
  # pivot_longer(cols = c("12-month Survival", "12-month % Dead", "12-month % Missing"), names_to = "12 month", values_to = "12 mo percentage") %>% 
  select(Year, Date, `Site ID`, Deliverable, `1 month`, `1 mo percentage`) %>% 
  mutate_at(c("1 month"), as.factor) %>% 
  mutate(`Site ID` = factor(`Site ID`, levels = mixedsort(levels(`Site ID`))))


# OneMoSurvivalAPAL2$`Site ID` <- factor(OneMoSurvivalAPAL2$`Site ID`, levels = mixedsort(levels(OneMoSurvivalAPAL2$`Site ID`)))




# %>% 
#   filter(between(Date, as.Date("2023-07-01"), as.Date("2024-02-01")))

```


## APAL One Month Survival Plot
Plotting One Month Survival Using a stacked bar graph
```{r, oneMonthSurvivalPlotAPAL, include = TRUE}

apal1MoSurvivalPlot1<- ggplot(OneMoSurvivalAPAL2, aes(x = `Site ID`, y = `1 mo percentage`, fill = `1 month`, color = I('black')))+
  geom_bar(position = "stack", stat = 'identity', width = 1)+
  ggtitle(expression("1 Month Percent Survival of "*italic("Acropora palmata")*"  Outplants"))+
  xlab("")+
  ylab("% Survival")+
  scale_fill_manual(values = myColors)

apal1MoSurvivalPlot <- apal1MoSurvivalPlot1 + theme(
    # Title
    plot.title = element_text(size = 20, face = "bold"),
    # X Axis
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_blank(),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "none",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))




ggsave("../figures/oneMonthSurvivalAPAL.png", plot =apal1MoSurvivalPlot, width = 20, height = 15, units = 'in', dpi = 600)



```


# APAL 12 Month Monitoring

## APAL 12mo monitoring

```{r, 12MonthDataWranglingAPAL, include = TRUE}

twelveMoSurvivalAPAL2 <- outplantAPAL %>% 
  mutate_at(c("1-month Survival", "1-month % Dead", "1-month % Missing", "12-month Survival", "12-month % Dead", "12-month % Missing"), as.double) %>% 
  # pivot_longer(cols = c("1-month Survival", "1-month % Dead", "1-month % Missing"), names_to = "1 month", values_to = "1 mo percentage") %>%
  pivot_longer(cols = c("12-month Survival", "12-month % Dead", "12-month % Missing"), names_to = "12 month", values_to = "12 mo percentage") %>% 
  select(Year, Date, `Site ID`, Deliverable, `12 month`, `12 mo percentage`) %>% 
    filter(!`Site ID` %in% c("T_AP_23", "T_AP_24", "T_AP_25", "T_AP_26")) %>% 
  na.omit() %>% 
  mutate_at(c("12 month"), as.factor) %>% 
    mutate(`Site ID` = factor(`Site ID`, levels = mixedsort(levels(`Site ID`))))
# %>% 
#   filter(between(Date, as.Date("2023-07-01"), as.Date("2024-02-01")))
 


```

## APAL 12 Month Survival Plot
```{r, 12MonthSurvivalPlotAPAL, include = TRUE}
apal12MoSurvivalPlot1<- ggplot(twelveMoSurvivalAPAL2, aes(x = `Site ID`, y = `12 mo percentage`, fill = `12 month`, color = I('black')))+
  geom_bar(position = "stack", stat = 'identity', width = 1)+
ggtitle(expression("12 Month Percent Survival of "*italic("Acropora palmata")*" Outplants"))+
  xlab("")+
  ylab("Percent Survival")+
  scale_fill_manual(values = myColors)



apal12MoSurvivalPlot <- apal12MoSurvivalPlot1 + theme(
    # Title
    plot.title = element_text(size = 20, face = "bold"),
    # X Axis
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_blank(),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "none",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))

ggsave("../figures/12MonthSurvivalPlotAPAL.png", plot =apal12MoSurvivalPlot, width = 20, height = 15, units = 'in', dpi = 600)



```




# ACER One Month Monitoring
## ACER One Month Data Wrangling

```{r, oneMonthDataWranglingACER, include = TRUE}


myColors <- c("red", "yellow", "green")
OneMoSurvivalACER2 <- outplantACER %>%
  filter(!is.na(`1-month Survival`)) %>% 
  mutate_at(c("1-month Survival", "1-month % Dead", "1-month % Missing"), as.numeric) %>%
  pivot_longer(cols = c("1-month Survival", "1-month % Dead", "1-month % Missing"), names_to = "1 month", values_to = "1 mo percentage") %>%
  select(Year, Date, `Site ID`, Deliverable, `1 month`, `1 mo percentage`) %>% 
  mutate_at(c("1 month"), as.factor)

```
## ACER One Month Survival Plot
```{r, oneMonthSurvivalPlotACER, include = TRUE}

acer1MoSurvivalPlot1<- ggplot(OneMoSurvivalACER2, aes(x = `Site ID`, y = `1 mo percentage`, fill = `1 month`, color = I('black')))+
  geom_bar(position = "stack", stat = 'identity', width = 1)+
ggtitle(expression("1 Month Percent Survival of "*italic("Acropora cervicornis")* " Outplants"))+
  xlab("")+
  ylab("Percent Survival")+
  scale_fill_manual(values = myColors)

acer1MoSurvivalPlot <- acer1MoSurvivalPlot1 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(size = 30, angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 40, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 40, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 30, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_blank(),
    legend.text = element_text(size = 40, face = "bold", colour = "black"),
    legend.position = "none",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))






ggsave("../figures/oneMonthSurvivalACER.png", plot = acer1MoSurvivalPlot, width = 20, height = 15, units = 'in', dpi = 600)


```



# ACER 12 Month Monitoring

## ACER 12mo monitoring

```{r, 12MonthDataWranglingACER, include = TRUE}

twelveMoSurvivalACER2 <- outplantACER %>% 
    filter(!is.na(`12-month Survival`)) %>%
  mutate_at(c("1-month Survival", "1-month % Dead", "1-month % Missing", "12-month Survival", "12-month % Dead", "12-month % Missing"), as.double) %>% 
  pivot_longer(cols = c("12-month Survival", "12-month % Dead", "12-month % Missing"), names_to = "12 month", values_to = "12 mo percentage") %>% 
  select(Year, Date, `Site ID`, Deliverable, `12 month`, `12 mo percentage`) %>% 
  mutate_at(c("12 month"), as.factor)

```



## ACER 12 Month Survival Plot
```{r, 12MonthSurvivalPlotACER, include = TRUE}

acer12MoSurvivalPlot1<- ggplot(twelveMoSurvivalACER2, aes(x = `Site ID`, y = `12 mo percentage`, fill = `12 month`, color = I('black')))+
  geom_bar(position = "stack", stat = 'identity', width = 1)+
  ggtitle(expression("12 Month Percent Survival of "*italic("Acropora cervicornis")*"  Outplants"))+
  xlab("Site")+
  ylab("Percent Survival")+
  scale_fill_manual(values = myColors)



acer12MoSurvivalPlot <- acer12MoSurvivalPlot1 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(size = 40, angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))



ggsave("../figures/12MonthSurvivalPlotACER.png", plot = acer12MoSurvivalPlot, width = 20, height = 15, units = 'in', dpi = 600)
```





## MSVS One Month Data Wrangling
```{r, oneMonthDataWranglingMSVS, include = TRUE}

myColors <- c("red", "yellow", "green")
OneMoSurvivalMSVS2 <- outplantMsvs %>% 
  select(-c(species, plugsPerSpecies)) %>% 
  distinct() %>%
  filter(!is.na(`1-month Survival`)) %>% 
  # mutate_at(c("1-month Survival", "1-month % Dead", "1-month % Missing", "12-month Survival", "12-month % Dead", "12-month % Missing"), as.double) %>%
    pivot_longer(cols = c("1-month Survival", "1-month % Dead", "1-month % Missing"), names_to = "1 month", values_to = "1 mo percentage") %>%
  # pivot_longer(cols = c("12-month Survival", "12-month % Dead", "12-month % Missing"), names_to = "12 month", values_to = "12 mo percentage") %>% 
  select(Date, `Site ID`, Deliverable, `1 month`, `1 mo percentage`) %>% 
  mutate_at(c("1 month"), as.factor) %>% 
  mutate(`Site ID` = factor(`Site ID`, levels = mixedsort(levels(`Site ID`))))


# OneMoSurvivalMSVS2$`Site ID` <- factor(OneMoSurvivalMSVS2$`Site ID`, levels = mixedsort(levels(OneMoSurvivalMSVS2$`Site ID`)))




# %>% 
#   filter(between(Date, as.Date("2023-07-01"), as.Date("2024-02-01")))

```





## MSVS One Month Survival Plot
Plotting One Month Survival Using a stacked bar graph
```{r, oneMonthSurvivalPlotMSVS, include = TRUE}

msvs1MoSurvivalPlot1<- ggplot(OneMoSurvivalMSVS2, aes(x = `Site ID`, y = `1 mo percentage`, fill = `1 month`, color = I('black')))+
  geom_bar(position = "stack", stat = 'identity', width = 1)+
  ggtitle(expression("1 Month Percent Survival of Massive Species Outplants"))+
  xlab("")+
  ylab("% Survival")+
  scale_fill_manual(values = myColors)

msvs1MoSurvivalPlot <- msvs1MoSurvivalPlot1 + theme(
    # Title
    plot.title = element_text(size = 20, face = "bold"),
    # X Axis
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_blank(),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "right",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))




ggsave("../figures/oneMonthSurvivalMSVS.png", plot =msvs1MoSurvivalPlot, width = 20, height = 15, units = 'in', dpi = 600)



```



# MSVS 12 Month Monitoring

## MSVS 12mo monitoring

```{r, 12MonthDataWranglingMSVS, include = TRUE}

twelveMoSurvivalMSVS2 <- outplantMsvs %>% 
  select(-c(species, plugsPerSpecies)) %>% 
  distinct() %>%
  filter(!is.na(`12-month Survival`)) %>% 
mutate(`12-month % Dead` = if_else(is.na(`12-month % Dead`), 0, `12-month % Dead`)) %>% 
  mutate(`12-month % Missing` = if_else(is.na(`12-month % Missing`), 0, `12-month % Missing`)) %>%
  pivot_longer(cols = c("12-month Survival", "12-month % Dead", "12-month % Missing"), names_to = "12 month", values_to = "12 mo percentage") %>% 
  select(Date, `Site ID`, Deliverable, `12 month`, `12 mo percentage`) %>% 
  mutate_at(c("12 month"), as.factor) %>% 
    mutate(`Site ID` = factor(`Site ID`, levels = mixedsort(levels(`Site ID`))))
# %>% 
#   filter(between(Date, as.Date("2023-07-01"), as.Date("2024-02-01")))
 


```

## MSVS 12 Month Survival Plot
```{r, 12MonthSurvivalPlotMSVS, include = TRUE}
msvs12MoSurvivalPlot1<- ggplot(twelveMoSurvivalMSVS2, aes(x = `Site ID`, y = `12 mo percentage`, fill = `12 month`, color = I('black')))+
  geom_bar(position = "stack", stat = 'identity', width = 1)+
ggtitle(expression("12 Month Percent Survival of Massive Species Outplants"))+
  xlab("")+
  ylab("Percent Survival")+
  scale_fill_manual(values = myColors)



msvs12MoSurvivalPlot <- msvs12MoSurvivalPlot1 + theme(
    # Title
    plot.title = element_text(size = 20, face = "bold"),
    # X Axis
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_blank(),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "right",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))

ggsave("../figures/12MonthSurvivalPlotMSVS.png", plot =msvs12MoSurvivalPlot, width = 20, height = 15, units = 'in', dpi = 600)







```





#Combined Plot
Combining the APAL and Massive Survival Plots 

```{r, comboPlot, include = TRUE}



comboSurvivalPlot <- apal1MoSurvivalPlot+msvs1MoSurvivalPlot+apal12MoSurvivalPlot+msvs12MoSurvivalPlot


ggsave("../figures/comboSurvivalplot.png", plot = comboSurvivalPlot, width = 20, height = 15, units = 'in', dpi = 600)


```





