---
title: "Analysis of Herbivory Assay: Herbivore Community Composition"
author: "Ian Combs -- icombs@mote.org"
date created: "07/30/2021"
last modified: "01/08/2022"
output:
  html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---
```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
library(magrittr)
```

### version: `r Sys.Date() %>% format(format="%B %d, %Y")`
###
***
This is the analysis pipeline to analyze the community composition of herbivores predating upon the herbivory assays. Six experimental 10m x 10m plots were established in Year 1 and 3 additional experimental plots were estabished in Y2. Two herbivory assays were deployed at each site. Each assay consisted for 5 blades of **Thalassia testudinum** set in clothespins threaded onto galvanized wire at 6in. intervals. The assays were deployed for 2 hours and filmed using a GoPro camera. The videos were analyzed to determine which species of herbivorous herb were predating upon the herbivory assays. Only herb that actually ate seagrass were noted, the blades that were predated upon were noted to the best of the reviewers ability, and, if applicable, the phase of the herbivore was noted (terminal, initial, etc).


***

### All analyses performed with R version `r getRversion()`


# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 

```{r,packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "officer", "ggpubr", "rcompanion", "RColorBrewer", "patchwork", "magrittr", "reshape2", "stringr", "plyr", "dplyr", "flextable", "tidyr", "tibble", "vegan", "forcats", "readr", "metafolio", "FSA", "viridis", "paletteer", 'tidyverse', "ecodist", "glue", "readxl")
pacman::p_load_gh("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")



```



# Loading data
***
Data generated from reviewing these videos are stored in an excel spreadsheet found in the ../data folder. And loaded in using the `read.xls()` from the package `gdata`.  

```{r, load and subset, include = TRUE, results = "hide"}
herb1 <- read_excel("../data/herbivoryVideoDataSheet.xlsx", sheet = 1)
herb1 <- herb1 %>% mutate(
    date = as.Date(date),
    site = as.factor(site),
    assay = as.factor(assay),
    reviewer1SpeciesCm = as.factor(reviewer1SpeciesCm),
    reviewer1SpeciesGn = as.factor(reviewer1SpeciesGn),
    reviewer1SpeciesSc = as.factor(reviewer1SpeciesSc),
    reviewer1FullSciName = as.factor(reviewer1FullSciName)) %>% 
    dplyr::rename("species" = "reviewer1FullSciName", "count" = "reviewer1Count")




# Creating date strings for the two separate years (Y1 and Y2)
Y1Dates <- herb1 %>%
  filter(site %in% c("IC-C1", "IC-Z2", "IC-U2")) %>%
  pull(date) %>%
  unique()


Y2Dates <- herb1 %>%
  filter(site %in% c("IC-C3", "IC-Z3", "IC-U3")) %>%
  pull(date) %>%
  unique()


# separating columns by Y1 and Y2 to add some additional column names, and then merging them again. 
# Filter and create separate data frames for each timePoint
# Y1
herbY1_T0 <- herb1 %>%
  filter(!site %in% c("IC-C3", "IC-Z3", "IC-U3")) %>%
  filter(date < as.Date("2021-10-16")) %>%
  mutate(timePoint = "T0", year = "Y1")

herbY1_T6 <- herb1 %>%
  filter(!site %in% c("IC-C3", "IC-Z3", "IC-U3")) %>%
  filter(date >= as.Date("2021-06-10") & date <= as.Date("2021-06-17")) %>%
  mutate(timePoint = "T6", year = "Y1")


herbY1_T12 <- herb1 %>%
  filter(!site %in% c("IC-C3", "IC-Z3", "IC-U3")) %>%
  filter(date >= as.Date("2022-01-20") & date <= as.Date("2022-01-30")) %>%
  mutate(timePoint = "T12", year = "Y1")

herbY1_T18 <- herb1 %>%
  filter(!site %in% c("IC-C3", "IC-Z3", "IC-U3")) %>%
  filter(date >= as.Date("2022-07-10")) %>% 
  mutate(timePoint = "T18", year = "Y1")

# Y2

herbY2_T0 <- herb1 %>%
  filter(!site %in% c("IC-C1", "IC-Z2", "IC-U2")) %>%
  filter(date >= as.Date("2020-10-16") & date <= as.Date("2021-06-16")) %>%
  mutate(timePoint = "T0", year = "Y2")

herbY2_T6 <- herb1 %>%
  filter(!site %in% c("IC-C1", "IC-Z2", "IC-U2")) %>%
  filter(date >= as.Date("2022-01-24") & date <= as.Date("2022-01-28")) %>%
  mutate(timePoint = "T6", year = "Y2")

herbY2_T12 <- herb1 %>%
  filter(!site %in% c("IC-C1", "IC-Z2", "IC-U2")) %>%
  filter(date >= as.Date("2022-07-09")) %>% 
  mutate(timePoint = "T12", year = "Y2")










# Merge all data frames into one
herbCombo <- bind_rows(herbY1_T0, herbY1_T6, herbY1_T12, herbY1_T18, herbY2_T0, herbY2_T6, herbY2_T12)

# Convert timePoint to factor with specified levels
herbCombo$timePoint <- factor(herbCombo$timePoint, levels = c("T0", "T6", "T12", "T18"))


herbCombo <- herbCombo %>% 
  mutate(treatment = if_else(site %in% c("IC-C2", "IC-Z1", "IC-U1"), "Control", "Outplanted")) %>% 
  mutate(reef = if_else(site %in% c("IC-C1", "IC-C2", "IC-C3"), "Site C",
                        if_else(site %in% c("IC-U1", "IC-U2", "IC-U3"), "Site U", "Site Z"))) %>% 
  mutate_at(vars(treatment, reef, year, timePoint), as.factor) %>% 
  filter(!timePoint == "T18") %>% 
  droplevels()



# selecting only the columns we want to work with and calculating abundance as a proportion. If you want a percentage just multiply by 100 but that makes certain tests and transformations difficult. 

herb <- herbCombo %>% select(date, site, treatment, reef, year, timePoint, species, count) %>%
  group_by(date, site, treatment, reef, year, timePoint, species) %>% 
  mutate(count = ifelse(is.na(count), 0, count)) %>%
  summarise(count = sum(count)) %>% 
  group_by(date, treatment, reef, year, timePoint, site) %>% 
  mutate(totalObservations = sum(count, na.rm = TRUE)) %>% 
  mutate(abundance = count/totalObservations)

# 
# herbAbundTreat <- herbCombo %>% 
#   group_by(treatment) %>% 
#   mutate(totalObservations = sum(count, na.rm = TRUE)) %>% 
#   summarise(mean(abundance))

# 
# herbAbundReef <- herb %>% 
#   group_by(reef) %>% 
#   summarise(mean(abundance))



```
# Normality Tests
***
Here we are exploring the distribution of the data, if the data is not normally distributed we will try to transform the data to meet a normal distribution.  

```{r, normalityTest, include = TRUE}

#Testing normality of abundance and count
hist(herb$count)
shapiro.test(herb$count)


hist(asin(sqrt(herb$count)))
shapiro.test(asin(sqrt(herb$count)))

```

<br><br>


# PERMANOVA 
***
Because the data is not normally distributed and transformations could not achieve normality, we are conducting a PERMANOVA using the package `vegan` to look for any differences in abundance between treatments. 

```{r, dissimilarityMatrix, include = TRUE}
herbMat <- herb %>% 
  select(-c(abundance, totalObservations)) %>%
  tidyr::unite(sample, 1:6) %>% 
  spread(species,count) %>%
  replace(is.na(.), 0) %>%
  # mutate_at(c('sample'), as.character) %>% 
  column_to_rownames(var = 'sample')

dist_matrix <- as.matrix(bcdist(herbMat),byrow = TRUE)

dist_tbl <- dist_matrix %>% 
        as_tibble(rownames = 'samples')

samples <- dist_tbl %>% 
  select(samples)


 samples_lookup <- dist_tbl %>% 
   select(samples) %>% 
   separate(col = samples, into = c("date", "site", "treatment", "reef","year", "timePoint"), sep = "_", remove = FALSE)

dist_tbl %>% 
  pivot_longer(cols = -samples, names_to = "b", values_to = "distances") %>% 
  pivot_wider(names_from = "b", values_from = "distances")

dist_matrix <- dist_tbl %>%
  pivot_longer(cols=-samples, names_to="b", values_to="distances") %>%
  inner_join(., samples_lookup, by="samples") %>%
  inner_join(., samples_lookup, by=c("b" = "samples")) %>%
  # filter(day.x %in% days_wanted & day.y %in% days_wanted) %>%
  select(samples, b, distances) %>%
  pivot_wider(names_from="b", values_from="distances") %>%
  select(-samples) %>%
  as.dist()



```
running a PCOA 

```{r, PCoA, include = TRUE}


pcoa <- cmdscale(dist_matrix, k = 2, eig = TRUE, add = TRUE) # add k = 3, k = 5 to add columns, default is k=2
positions <- pcoa$points
colnames(positions) <- c("pcoa1", 'pcoa2')
# calculating the amount of variation explained by each dimension
percentExplained <- 100 * (pcoa$eig / sum(pcoa$eig))

#rounding and trimming to add to labels
RoundPercentExplained <- format(round(percentExplained[1:2], digits = 1), nsmall = 1, trim = TRUE)

# creating label vector
labs <- c(glue("PCo 1 ({RoundPercentExplained[1]}%)"),
          glue("PCo 2 ({RoundPercentExplained[2]}%)"))



treatmentCentroid <- positions %>% as_tibble(rownames='samples') %>%
  inner_join(., samples_lookup, by = 'samples') %>% 
  group_by(treatment) %>% 
  summarize(axis1 = mean(pcoa1),
            axis2 = mean(pcoa2))



#
pcoaTbl <- positions %>% as_tibble(rownames='samples') %>%
  inner_join(., samples_lookup, by = 'samples') 

pcoaPlot <- ggplot() +
  geom_point(data = pcoaTbl, mapping = aes(x = pcoa1, y = pcoa2, color = treatment, shape = timePoint), size = 3) +
  geom_point(data = treatmentCentroid, mapping = aes(x = axis1, y = axis2, color = treatment), shape = 15, size = 5) +
  labs(x = labs[1], y = labs[2])


pcoaPlot <- ggplot(pcoaTbl, aes(x = pcoa1, y = pcoa2, color = treatment, fill = treatment)) +
  stat_ellipse(geom = 'polygon', alpha = 0.2, level = 0.75, show.legend = FALSE) +
  geom_point(aes(x = pcoa1, y = pcoa2, shape = timePoint), alpha = 1, size = 3) +
  labs(x = labs[1], y = labs[2])



```


```{r, NMDS, include = TRUE}

set.seed(3008)
nmds <- metaMDS(dist_matrix)

scores(nmds) %>% 
  as_tibble(rownames='samples') %>% 
  inner_join(., samples_lookup, by = 'samples') %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = treatment))+
  geom_point(size = 3)
  




```
```{r, ecologicalDistances, include = TRUE}

ecoMat <- as.matrix(herbMat)
samples_lookup <- herbMat %>%
  rownames_to_column(var = 'samples') %>% 
   select(samples) %>% 
   separate(col = samples, into = c("date", "site", "treatment", "reef","year", "timePoint"), sep = "_", remove = FALSE)


set.seed(3008)
dist <- vegdist(ecoMat, method = 'bray')
nmds <- metaMDS(dist)

nmds <- scores(nmds) %>% 
  as_tibble(rownames = 'samples') %>%
  inner_join(., samples_lookup, by = 'samples') %>% 
  ggplot(aes(x = NMDS1, y = NMDS2,color = treatment)) +
  stat_ellipse(geom = 'polygon', alpha = 0.2, show.legend = FALSE) +
  geom_point(size = 3) 








```
```{r, PERMANOVA, include = TRUE, message = FALSE}

set.seed(9999)

 samplesLookupPerm <- herbMat %>%
   rownames_to_column(var = "samples") %>% 
   select(samples) %>% 
   separate(col = samples, into = c("date", "site", "treatment", "reef","year", "timePoint"), sep = "_", remove = FALSE) %>% 
   column_to_rownames(var = "samples") %>% 
   mutate_at(c("date", "site", "treatment", "reef","year", "timePoint"), as.factor)
 
vpiDist <- vegdist(herbMat, method = 'bray')

distPerm <- adonis2(herbMat ~ treatment*timePoint*reef, data = samplesLookupPerm, permutations = 9999, method = 'bray')
distPerm




```
```{r}
distPermTab <- distPerm %>%
  # add_column("Test",)
  as_tibble(rownames = "Comparison") %>%
  # mutate(Comparison, gsub(pattern = "timePoint", replacement = "Time Point", pattern = "year", replacement = "Year", pattern = "reef", replacement = "Reef")) %>% 
  rename(`Pseudo.F` = `F`, `p.value` = `Pr(>F)`) %>% 
  select(Comparison, Df, `Pseudo.F`, `p.value`) %>% 
  mutate_if(is.numeric, round, digits = 3) %>%
    mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
  mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>% 
  mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–") %>%
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Df", value = as_paragraph("Df")) %>% 
  flextable::compose(part = "header", j = "Pseudo.F", value = as_paragraph("Pseudo-F")) %>%
  flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
  autofit() %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  colformat_num(j = 'Df', digits = 2) %>% 
  colformat_num(j = "Pseudo.F", digits = 2) %>%
  colformat_num(j = "p.value", digits = 4, na_str = "ns") %>% 
  align_nottext_col(align = "center", header = TRUE, footer = TRUE) %>% 
  align(align = "center", j = "p.value")

distDoc = read_docx()
distDoc = body_add_flextable(distDoc, value = distPermTab)
print(distDoc, target = "../tables/distTableherb.docx")
distPermTab

```

`````{r, plotting, include = TRUE, eval = FALSE}
myColors <- paletteer_c("pals::ocean.balance", 18, direction = 1 )


# Making a box plot for all Y1 data

boxPlotData <- herbMat %>% 
   rownames_to_column(var = "samples") %>% 
    separate(col = samples, into = c("date", "site", "treatment", "reef","year", "timePoint"), sep = "_", remove = FALSE) %>% 
  select(-c(samples)) %>% 
  gather(class, count, 7:18, factor_key = TRUE) %>% 
  mutate(class = fct_reorder(class, count, .desc = TRUE)) %>%   as_tibble() %>% 
  mutate_at(c("date", "site", "treatment", "reef","year", "timePoint"), as.factor) %>% 
  group_by(reef, treatment, class) %>% 
  summarise(sumCount = sum(count)) %>% 
  arrange(desc(sumCount)) %>% 
  group_by(reef, treatment) %>% 
  mutate(totalObservations = sum(sumCount)) %>% 
  group_by(reef, treatment, class) %>% 
  mutate(abundance = (sumCount/totalObservations)*100) %>% 
  # filter(abundance > 1) %>% 
  arrange(desc(abundance))


herbStack1 <- ggplot(boxPlotData, aes(x = reef, y = abundance)) +
                      geom_bar(aes(fill = class), position = position_stack(reverse = TRUE), stat = 'identity', color = 'black', alpha = 0.8) +
                      ylim(0,100) +
                      labs(y = "Relative Abundance (%)", fill = "Species", title = "Relative Species Abundance")+
                      scale_fill_manual(values = myColors)+
                      facet_wrap(.~ treatment, scales = 'free_x')




herbStack <- herbStack1+theme(
    # Title
    plot.title = element_text(size = 30, face = "bold"), 
    # X Axis
    axis.text.x = element_text(size = 30, colour = "black", vjust = 0.5, hjust = 0.5, face= "bold"), 
    axis.title.x = element_blank(),
    # Y Axis
    axis.title.y = element_text(size = 30, face = "bold"), 
    axis.text.y = element_text(colour = "black", size = 40, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"), 
    axis.ticks = element_line(color="black"), 
    # Legend
    legend.title = element_text(size = 25, face = "bold"), 
    legend.text = element_text(size = 15, face = "bold", colour = "black"),
    legend.position = "right",
    # Grid and Background
    panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
    panel.background = element_rect(fill = '#F5F5F5'), 
    # Text
    text = element_text(size = 30, color="black")) 









herbStack









ggsave("../figures/herbStackY1.png", plot = herbStack, width = 20, height = 15, units = 'in', dpi = 600)


```

```{r, histogram, include = TRUE}

herbHist <- ggplot(herb, aes(x = class, y = count)) +
                      geom_boxplot(aes(fill = class), position = position_stack(), stat = 'identity', color = 'black', alpha = 0.8) +
                      ylim(0,100) +
                      labs(y = "Relative Abundance (%)", fill = "Class", title = "Relative Species Abundance > 5%")+
                      scale_fill_manual(values = myColors)+
                      facet_wrap(treatment ~ year, scales = 'free')









herbHist1 <- ggplot(herb, aes(x = species, y = count))  +
  geom_bar(aes(fill = treatment), position = 'dodge', stat = 'identity')+
  facet_grid(year*timePoint ~ reef, scales = "free_x")












herbHist <- herbHist1+theme(
    # Title
    plot.title = element_text(size = 30, face = "bold"), 
    # X Axis
    axis.text.x = element_text(size = 30, colour = "black", vjust = 0.5, hjust = 0.5, face= "bold"), 
    axis.title.x = element_blank(),
    # Y Axis
    axis.title.y = element_text(size = 30, face = "bold"), 
    axis.text.y = element_text(colour = "black", size = 40, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"), 
    axis.ticks = element_line(color="black"), 
    # Legend
    legend.title = element_text(size = 25, face = "bold"), 
    legend.text = element_text(size = 15, face = "bold", colour = "black"),
    legend.position = "right",
    # Grid and Background
    panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
    panel.background = element_rect(fill = '#F5F5F5'), 
    # Text
    text = element_text(size = 30, color="black")) 





```
