---
title: "Fish Video Transect Survey Analysis"
author: "Ian Combs -- icombs@mote.org"
date created: 01/13/2022
output:
  html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '3'
---


```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
library(magrittr)
```

### version: `r Sys.Date() %>% format(format="%B %d, %Y")`
###
***
This is the analysis pipeline to analyze the community composition from fish video transects. Six experimental 10m x 10m plots were established in August 2020. Three more experimental 10m x 10m plots were established in June of 2021. Three transects of video were captured per site three times 1 month post-outplant and again, three more times 12 months post-outplant. Each transect was swam from the ledge to the hardpan, and 5 minutes elapsed between transects to allow for fish communities to return post diver disturbance. Still frames were pulled from each video every 3 seconds. Using the open source program **Ffmpeg**, the protocol for which can be found in the ../protocols folder within this repository. Fish were identified to the lowest taxonomic level possible, if applicable, the phase of the fish species was noted (terminal, initial, etc).

***

### All analyses performed with R version `r getRversion()`


# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some may be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 

```{r,packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "officer", "ggpubr", "rcompanion", "RColorBrewer", "patchwork", "magrittr", "reshape2", "stringr", "plyr", "dplyr", "flextable", "tidyr", "tibble", "vegan", "forcats", "readr", "metafolio", "FSA", "viridis", "paletteer", 'tidyverse', "ecodist", "glue")
pacman::p_load_gh("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")

```


# Loading data
***
Data generated from reviewing these videos and stills are stored in an excel spreadsheet found in the ../data folder. And loaded in using the `readxlsx()` from the package `readxl`. Note: we are also loading the `officer` package which has a similarly named function, so we are explicitly calling the `read_xlsx()` function from `readxl` using `::` notation.   


```{r, load and subset, include = TRUE, results = "hide"}

fish1 <- readxl::read_xlsx("../data/fishSurveyData.xlsx", sheet = 1, col_names = TRUE, na = c("", "N/A", "NA", "NA NA", "UNK","UNK UNK", "N/A N/A","NA  NA"))

#turning a bunch of the columns into factors, and renaming reviewer1FullSCiName to species
fish1 <- fish1 %>%
  mutate(across(c(site, transect, frame, reviewer1SpeciesCm, 
                   reviewer1SpeciesGn, reviewer1SpeciesSc, reviewer1FullSciName,
                   phase), as.factor), 
         date = as.Date(date)) %>% 
  dplyr::rename("species" = "reviewer1FullSciName")



# Creating date strings for the two separate years (Y1 and Y2)
Y1Dates <- fish1 %>%
  filter(site %in% c("IC-C1", "IC-Z2", "IC-U2")) %>%
  pull(date) %>%
  unique()


Y2Dates <- fish1 %>%
  filter(site %in% c("IC-C3", "IC-Z3", "IC-U3")) %>%
  pull(date) %>%
  unique()


# separating columns by Y1 and Y2 to add some additional column names, and then merging them again. 

fishY1 <- fish1 %>% 
  filter(date == Y1Dates) %>%
  mutate(treatment = if_else(site %in% c("IC-C2", "IC-Z1", "IC-U1"), "Control","Outplanted")) %>% 
  mutate(reef = if_else(site %in% c("IC-C1", "IC-C2", "IC-C3"), "Site C",
                   if_else(site %in% c("IC-U1", "IC-U2", "IC-U3"), "Site U", "Site Z"))) %>% 
  mutate(year = "Y1") %>% 
  mutate(timePoint =if_else(date < "2021-01-19", "T0", "T12")) %>% 
  droplevels() %>% 
  mutate_at(vars(treatment, reef, year, timePoint), as.factor)


fishY2 <- fish1 %>% 
  filter(date == Y2Dates) %>% 
  mutate(treatment = if_else(site %in% c("IC-C2", "IC-Z1", "IC-U1"), "Control","Outplanted")) %>% 
  mutate(reef = if_else(site %in% c("IC-C1", "IC-C2", "IC-C3"), "Site C",
                   if_else(site %in% c("IC-U1", "IC-U2", "IC-U3"), "Site U", "Site Z"))) %>% 
  mutate(year = "Y2") %>% 
  mutate(timePoint =if_else(date < "2021-07-24", "T0", "T12")) %>% 
  droplevels() %>% 
  mutate_at(vars(treatment, reef, year, timePoint), as.factor)

fishCombo <- bind_rows(fishY1, fishY2)








# selecting only the columns we want to work with and calculating abundance as a proportion. If you want a percentage just multiply by 100 but that makes certain tests and transformations difficult. 

fish2 <- fishCombo %>% select(date, site, treatment, reef, year, timePoint, species, count) %>%
  group_by(date, site, treatment, reef, year, timePoint, species) %>% 
  mutate(count = ifelse(is.na(count), 0, count)) %>% 
  add_count() %>% 
  rename(tally = 'count') %>% 
  summarise(count = sum(tally), n = max(n), avgCount = ((count/n)))


fish3 <- fish2 %>% 
  group_by(date, site, treatment, reef, year, timePoint, species) %>% 
  select(-c(tally, n, count)) %>% 
  add_count() %>% 
  summarise(count = sum(avgCount), n = max(n), avgCount = ((count/n))) 

fish <- fish3 %>%
  group_by(date, site, treatment, reef, year, timePoint, species) %>% 
  group_by(date, site) %>% 
  mutate(totalObservations = sum(avgCount, na.rm = TRUE)) %>%
  group_by(date, site, species) %>% 
  mutate(abundance = avgCount/totalObservations)

```
# Normality Tests
***
Here we are exploring the distribution of the data, if the data is not normally distributed we will try to transform the data to meet a normal distribution.  

```{r, normalityTest, include = TRUE}

#Testing normality of abundance and count
hist(fish$abundance)
shapiro.test(fish$abundance)



hist(asin(sqrt(fish$abundance)))
shapiro.test(asin(sqrt(fish$abundance)))




```

# PERMANOVA 
***
Because the data is not normally distributed and transformations could not achieve normality, we are conducting a PERMANOVA using the package `vegan` to look for any differences in abundance between treatments. 

```{r, dissimilarityMatrix, include = TRUE}
fishMat <- fish %>% 
  select(-c(abundance, count,n, totalObservations, transect, frame)) %>%
  tidyr::unite(sample, 1:6) %>% 
  spread(species,avgCount) %>%
  replace(is.na(.), 0) %>%
  # mutate_at(c('sample'), as.character) %>% 
  column_to_rownames(var = 'sample')

dist_matrix <- as.matrix(bcdist(fishMat),byrow = TRUE)

dist_tbl <- dist_matrix %>% 
        as_tibble(rownames = 'samples')

samples <- dist_tbl %>% 
  select(samples)


 samples_lookup <- dist_tbl %>% 
   select(samples) %>% 
   separate(col = samples, into = c("date", "site", "reef", "treatment", "year", "timePoint"), sep = "_", remove = FALSE)

dist_tbl %>% 
  pivot_longer(cols = -samples, names_to = "b", values_to = "distances") %>% 
  pivot_wider(names_from = "b", values_from = "distances")

dist_matrix <- dist_tbl %>%
  pivot_longer(cols=-samples, names_to="b", values_to="distances") %>%
  inner_join(., samples_lookup, by="samples") %>%
  inner_join(., samples_lookup, by=c("b" = "samples")) %>%
  # filter(day.x %in% days_wanted & day.y %in% days_wanted) %>%
  select(samples, b, distances) %>%
  pivot_wider(names_from="b", values_from="distances") %>%
  select(-samples) %>%
  as.dist()



```
running a PCOA 

```{r, PCoA, include = TRUE}


pcoa <- cmdscale(dist_matrix, k = 2, eig = TRUE, add = TRUE) # add k = 3, k = 5 to add columns, default is k=2
positions <- pcoa$points
colnames(positions) <- c("pcoa1", 'pcoa2')
# calculating the amount of variation explained by each dimension
percentExplained <- 100 * (pcoa$eig / sum(pcoa$eig))

#rounding and trimming to add to labels
RoundPercentExplained <- format(round(percentExplained[1:2], digits = 1), nsmall = 1, trim = TRUE)

# creating label vector
labs <- c(glue("PCo 1 ({RoundPercentExplained[1]}%)"),
          glue("PCo 2 ({RoundPercentExplained[2]}%)"))



treatmentCentroid <- positions %>% as_tibble(rownames='samples') %>%
  inner_join(., samples_lookup, by = 'samples') %>% 
  group_by(treatment) %>% 
  summarize(axis1 = mean(pcoa1),
            axis2 = mean(pcoa2))



#
pcoaTbl <- positions %>% as_tibble(rownames='samples') %>%
  inner_join(., samples_lookup, by = 'samples') 

pcoaPlot <- ggplot() +
  geom_point(data = pcoaTbl, mapping = aes(x = pcoa1, y = pcoa2, color = treatment, shape = timePoint), size = 3) +
  geom_point(data = treatmentCentroid, mapping = aes(x = axis1, y = axis2, color = treatment), shape = 15, size = 5) +
  labs(x = labs[1], y = labs[2])


pcoaPlot <- ggplot(pcoaTbl, aes(x = pcoa1, y = pcoa2, color = treatment, fill = treatment)) +
  stat_ellipse(geom = 'polygon', alpha = 0.2, level = 0.75, show.legend = FALSE) +
  geom_point(aes(x = pcoa1, y = pcoa2, shape = timePoint), alpha = 1, size = 3) +
  labs(x = labs[1], y = labs[2])



```


```{r, NMDS, include = TRUE}

set.seed(3008)
nmds <- metaMDS(dist_matrix)

scores(nmds) %>% 
  as_tibble(rownames='samples') %>% 
  inner_join(., samples_lookup, by = 'samples') %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = treatment))+
  geom_point(size = 3)
  




```
```{r, ecologicalDistances, include = TRUE}

ecoMat <- as.matrix(fishMat)
samples_lookup <- fishMat %>%
  rownames_to_column(var = 'samples') %>% 
   select(samples) %>% 
   separate(col = samples, into = c("date", "timePoint", "year", "site", "treatment", "reef"), sep = "_", remove = FALSE)


set.seed(3008)
dist <- vegdist(ecoMat, method = 'bray')
nmds <- metaMDS(dist)

scores(nmds) %>% 
  as_tibble(rownames = 'samples') %>%
  inner_join(., samples_lookup, by = 'samples') %>% 
  ggplot(aes(x = NMDS1, y = NMDS2,color = treatment)) +
  stat_ellipse(geom = 'polygon', alpha = 0.2, show.legend = FALSE) +
  geom_point(size = 3) 








```
```{r, PERMANOVA, include = TRUE, message = FALSE}

set.seed(9999)

 samplesLookupPerm <- fishMat %>%
   rownames_to_column(var = "samples") %>% 
   select(samples) %>% 
   separate(col = samples, into = c("date", "site", "reef", "treatment", "year", "timePoint"), sep = "_", remove = FALSE) %>% 
   column_to_rownames(var = "samples") %>% 
   mutate_at(c("date", "timePoint", "year", "site", "treatment", "reef"), as.factor)
 
vpiDist <- vegdist(fishMat, method = 'bray')

distPerm <- adonis2(fishMat ~ treatment*timePoint*reef, data = samplesLookupPerm, permutations = 9999, method = 'bray')
distPerm




```
```{r}
distPermTab <- distPerm %>%
  # add_column("Test",)
  as_tibble(rownames = "Comparison") %>%
  # mutate(Comparison, gsub(pattern = "timePoint", replacement = "Time Point", pattern = "year", replacement = "Year", pattern = "reef", replacement = "Reef")) %>% 
  rename(`Pseudo.F` = `F`, `p.value` = `Pr(>F)`) %>% 
  select(Comparison, Df, `Pseudo.F`, `p.value`) %>% 
  mutate_if(is.numeric, round, digits = 3) %>%
    mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
  mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>% 
  mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–") %>%
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Df", value = as_paragraph("Df")) %>% 
  flextable::compose(part = "header", j = "Pseudo.F", value = as_paragraph("Pseudo-F")) %>%
  flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
  autofit() %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  colformat_num(j = 'Df', digits = 2) %>% 
  colformat_num(j = "Pseudo.F", digits = 2) %>%
  colformat_num(j = "p.value", digits = 4, na_str = "ns") %>% 
  align_nottext_col(align = "center", header = TRUE, footer = TRUE) %>% 
  align(align = "center", j = "p.value")

distDoc = read_docx()
distDoc = body_add_flextable(distDoc, value = distPermTab)
print(distDoc, target = "../tables/distTableFish.docx")
distPermTab

```

`````{r, plotting, include = TRUE, eval = FALSE}
myColors <- paletteer_c("pals::ocean.balance", 21, direction = 1 )


# Making a box plot for all Y1 data

boxPlotDataY1 <- fishMat %>% 
   rownames_to_column(var = "samples") %>% 
    separate(col = samples, into = c("date", "site", "reef", "treatment", "year", "timePoint"), sep = "_", remove = FALSE) %>% 
  select(-c(samples)) %>% 
  gather(class, count, 7:88, factor_key = TRUE) %>% 
  as_tibble() %>% 
  mutate_at(c("date", "timePoint", "year", "site", "treatment", "reef", "class"), as.factor) %>% 
  group_by(year, reef, site, treatment, class) %>% 
  summarise(sumCount = sum(count)) %>% 
  arrange(desc(sumCount)) %>% 
  group_by(year, reef, site, treatment) %>% 
  mutate(totalObservations = sum(sumCount)) %>% 
  group_by(year, reef, site, treatment, class) %>% 
  mutate(abundance = (sumCount/totalObservations)*100) %>% 
  filter(abundance > 5) %>% 
  arrange(desc(abundance))


fishStack1Y1 <- ggplot(boxPlotDataY1, aes(x = site, y = abundance)) +
                      geom_bar(aes(fill = class), position = position_stack(), stat = 'identity', color = 'black', alpha = 0.8) +
                      ylim(0,100) +
                      labs(y = "Relative Abundance (%)", fill = "Class", title = "Relative Species Abundance > 5%")+
                      scale_fill_manual(values = myColors)+
                      facet_wrap(treatment ~ year, scales = 'free')




fishStackY1 <- fishStack1Y1+theme(
    # Title
    plot.title = element_text(size = 30, face = "bold"), 
    # X Axis
    axis.text.x = element_text(size = 30, colour = "black", vjust = 0.5, hjust = 0.5, face= "bold"), 
    axis.title.x = element_blank(),
    # Y Axis
    axis.title.y = element_text(size = 30, face = "bold"), 
    axis.text.y = element_text(colour = "black", size = 40, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"), 
    axis.ticks = element_line(color="black"), 
    # Legend
    legend.title = element_text(size = 25, face = "bold"), 
    legend.text = element_text(size = 15, face = "bold", colour = "black"),
    legend.position = "right",
    # Grid and Background
    panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
    panel.background = element_rect(fill = '#F5F5F5'), 
    # Text
    text = element_text(size = 30, color="black")) 









fishStackY1









ggsave("../figures/fishStackY1.png", plot = vpiStackY1, width = 20, height = 15, units = 'in', dpi = 600)


```

