---
title: "noaaVPI"
author: "Ian Combs -- icombs@mote.org"
output:
  html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
library(magrittr)
```

### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

<!-- this is where the DOI would go  [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3675991.svg)](https://doi.org/10.5281/zenodo.3675991)
-->


#### [GitHub repository](https://github.com/icombs2017/photogrammetryNOAA.git){target="_blank"}

###
***
This is the analysis pipeline for data generated from TagLab annotations of our 6 10x10m plots.


***
### All analyses performed with R verion `r getRversion()`
# Basic setup of R environment
***
## Loading required packages

For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 

```{r, load packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "officer", "ggpubr", "rcompanion", "RColorBrewer", "patchwork", "magrittr", "reshape2", "stringr", "plyr", "dplyr", "flextable", "tidyr", "tibble", "vegan", "forcats", "readr", "metafolio", "FSA", "viridis", "paletteer", 'tidyverse', "ecodist", "glue")
pacman::p_load_gh("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
```

<br><br>

# Loading Data
***

Data is exported from Viscore's VPI scriplet as a .csv file and named according to the file. We have multiple .csv's that we wish to compile so we created a function called `read_plus` that will read all the .csv's into one data frame witht he file name as a column. We will then parse out various pieces of the file name to provide us use ful columns e.g. date, site, etc. 
Since Viscore puts the "percent" column as a percentage sign (**%**) R cannot read it in  we are using `dplyr`'s `rename` function to change it to something R can read and that is easier to call and read when we are handling the data. I am also adding **site** and **time** columns to each subset, combining all subsets into one large dataframe using `bind_rows()`, adding a **treatment** column and using the `gsub()` to rename some.


```{r, loadData, include = TRUE, message = FALSE}


read_plus <- function(flnm) {
    read_csv(flnm) %>% 
        mutate(filename = flnm)
}

vpiData1 <-
    list.files(path = "../data/vpi", pattern = "*.csv", 
               full.names = T) %>% 
    map_df(~read_plus(.))


# Sort out data to have proper columns for all the info i want e.g. site, date etc

vpiData <- vpiData1 %>% 
      tidyr::separate(filename, into=c("Date", "Site"),extra = 'merge',sep = 23) %>%
      tidyr::separate(Date, into=c("remove", "Date"),sep = "../data/vpi/") %>% 
      tidyr::separate(Date, into=c("Date", "remove"),sep = 10) %>% 
      tidyr::separate(Site, into = c("Site", "delete"),sep = 5) %>% 
      dplyr::select(-c(remove, delete)) %>% 
      mutate_if(is.character, str_replace_all, pattern = "_", replacement = "-") %>%
      dplyr::rename(abundance = `%`) %>%
      tidyr::drop_na(abundance) %>% 
      mutate(Date = factor(Date), Site = factor(Site), timePoint = factor(timePoint), year = factor(year))

#Making some corrections in the naming

vpiData <- vpiData %>%
  mutate_all(~gsub("Crustose coralline algae natans", "Crustose coralline algae", .)) %>%
  mutate_all(~gsub("Cordata", "Chordata", .)) %>%
  mutate_all(~gsub("Rubble", "Bare Substrate", .)) %>%
  mutate_all(~gsub("Other|Review|None", NA, .))


vpiData<- na.omit(vpiData)
vpiData$class <- as.factor(vpiData$class)

vpiData$class <-  droplevels.factor(vpiData$class)

vpiData$treatment = if_else(vpiData$Site %in% c("IC-C2", "IC-Z1", "IC-U1"), "Control","Outplanted") 
vpiData$reef = if_else(vpiData$Site %in% c("IC-C1", "IC-C2", "IC-C3"), "SiteC",
                   if_else(vpiData$Site %in% c("IC-U1", "IC-U2", "IC-U3"), "SiteU", "SiteZ"))
vpiData$reef <- as.factor(vpiData$reef)

vpiData <- vpiData %>% 
      select(Date, timePoint, year, Site, treatment, reef, class, count, abundance) %>% 
      mutate(Date = as.Date(Date)) %>% 
    mutate(count = as.numeric(count), abundance = as.numeric(abundance)) %>% 
  mutate(across(where(is.character), factor))
  


head(vpiData)

```

Going from long to wide format (matrix) and also creating a separate dataframe with just the samples metadata. 

```{r, dissimilarityMatrix, include = TRUE}


vpiMat <- vpiData %>% 
  select(-c(abundance)) %>%
  tidyr::unite(sample, 1:6) %>% 
  spread(class,count) %>%
  replace(is.na(.), 0) %>%
  # mutate_at(c('sample'), as.character) %>% 
  column_to_rownames(var = 'sample')

dist_matrix <- as.matrix(bcdist(vpiMat),byrow = TRUE)


dist_tbl <- dist_matrix %>% 
        as_tibble(rownames = 'samples')

samples <- dist_tbl %>% 
  select(samples)


 samples_lookup <- dist_tbl %>% 
   select(samples) %>% 
   separate(col = samples, into = c("date", "timePoint", "year", "site", "treatment", "reef"), sep = "_", remove = FALSE)
 


  

dist_tbl %>% 
  pivot_longer(cols = -samples, names_to = "b", values_to = "distances") %>% 
  pivot_wider(names_from = "b", values_from = "distances")
 
 

dist_matrix <- dist_tbl %>%
  pivot_longer(cols=-samples, names_to="b", values_to="distances") %>%
  inner_join(., samples_lookup, by="samples") %>%
  inner_join(., samples_lookup, by=c("b" = "samples")) %>%
  # filter(day.x %in% days_wanted & day.y %in% days_wanted) %>%
  select(samples, b, distances) %>%
  pivot_wider(names_from="b", values_from="distances") %>%
  select(-samples) %>%
  as.dist()


```

Creating a PCoA using a distance matrix using curtis-bray distances


```{r, PCoA, include = TRUE}


pcoa <- cmdscale(dist_matrix, k = 2, eig = TRUE, add = TRUE) # add k = 3, k = 5 to add columns, default is k=2
positions <- pcoa$points
colnames(positions) <- c("pcoa1", 'pcoa2')
# calculating the amount of variation explained by each dimension
percentExplained <- 100 * (pcoa$eig / sum(pcoa$eig))

#rounding and trimming to add to labels
RoundPercentExplained <- format(round(percentExplained[1:2], digits = 1), nsmall = 1, trim = TRUE)

# creating label vector
labs <- c(glue("PCo 1 ({RoundPercentExplained[1]}%)"),
          glue("PCo 2 ({RoundPercentExplained[2]}%)"))



treatmentCentroid <- positions %>% as_tibble(rownames='samples') %>%
  inner_join(., samples_lookup, by = 'samples') %>% 
  group_by(treatment) %>% 
  dplyr::summarise(axis1 = mean(pcoa1),
            axis2 = mean(pcoa2))



#
pcoaTbl <- positions %>% as_tibble(rownames='samples') %>%
  inner_join(., samples_lookup, by = 'samples') 

pcoaPlot1 <- ggplot() +
  geom_point(data = pcoaTbl, mapping = aes(x = pcoa1, y = pcoa2, color = treatment, shape = timePoint), size = 3) +
  geom_point(data = treatmentCentroid, mapping = aes(x = axis1, y = axis2, color = treatment), shape = 15, size = 5) +
  labs(x = labs[1], y = labs[2])


pcoaPlot2 <- ggplot(pcoaTbl, aes(x = pcoa1, y = pcoa2, color = treatment, fill = treatment)) +
  stat_ellipse(geom = 'polygon', alpha = 0.2, level = 0.75, show.legend = FALSE) +
  geom_point(aes(x = pcoa1, y = pcoa2, shape = year), alpha = 1, size = 3) +
  ggtitle("Principle Coordinate Analysis of VPI Data")+ 
  labs(x = labs[1], y = labs[2])

pcoaPlot <- pcoaPlot2 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))

ggsave("../figures/pcoa.png", plot = pcoaPlot, width = 20, height = 15, units = 'in', dpi = 600)





tibble(pe = cumsum(percentExplained),
       axis = 1:length(percentExplained)) %>%
  ggplot(aes(x = axis, y = pe))+
  geom_line()+
  coord_cartesian(xlim = c(1,10))




```

Creating an NMDS using that same distance matrix


```{r, NMDS, include = TRUE}

set.seed(3008)
nmds <- metaMDS(dist_matrix)

scores(nmds) %>% 
  as_tibble(rownames='samples') %>% 
  inner_join(., samples_lookup, by = 'samples') %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = treatment))+
  geom_point()
  




```





```{r, ecologicalDistances, include = TRUE}

ecoMat <- as.matrix(vpiMat)
samples_lookup <- vpiMat %>%
  rownames_to_column(var = 'samples') %>% 
   select(samples) %>% 
   separate(col = samples, into = c("date", "timePoint", "year", "site", "treatment", "reef"), sep = "_", remove = FALSE)


set.seed(3008)
dist <- vegdist(ecoMat, method = 'bray')
nmds <- metaMDS(dist)

scores(nmds) %>% 
  as_tibble(rownames = 'samples') %>%
  inner_join(., samples_lookup, by = 'samples') %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, shape = treatment, color = reef))+
  geom_point()







```




<br><br>

# PERMANOVA
***

Running a PERMANOVA since the data does not meet the assumptions of normality. We are running the PERMANOVA on the dissimilarity matrix we created above but we are running it by removing Bare Substrate from our matrix.  

```{r, PERMANOVA, include = TRUE, message = FALSE}
# doing PERMANOVA on a dissimilarity matrix, 'old' code is commented below. 

set.seed(9999)

vpiMat1 <- vpiMat %>% 
  select(-c(`Bare Substrate`))



 samplesLookupPerm <- vpiMat1 %>% 
   rownames_to_column(var = "samples") %>% 
   select(samples) %>% 
   separate(col = samples, into = c("date", "timePoint", "year", "site", "treatment", "reef"), sep = "_", remove = FALSE) %>% 
   column_to_rownames(var = "samples")
 
vpiDist <- vegdist(vpiMat1, method = 'bray')

distPerm <- adonis2(vpiDist ~ treatment*year*timePoint, data = samplesLookupPerm, permutations = 9999, method = 'bray')

distPerm




```




<br><br>
# Output Table
***

Creating a publication-ready table from the output of our PERMANOVA above. This is mainly using the package `flextable`.  




```{r, distPermTab, include = TRUE}




distPermTab <- distPerm %>%
  as_tibble(rownames = "Comparison") %>%
  rename(`Pseudo.F` = `F`, `p.value` = `Pr(>F)`) %>% 
  select(Comparison, Df, `Pseudo.F`, `p.value`) %>% 
  mutate_if(is.numeric, round, digits = 3) %>%
    mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
  mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>% 
  mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–") %>%
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Df", value = as_paragraph("Df")) %>% 
  flextable::compose(part = "header", j = "Pseudo.F", value = as_paragraph("Pseudo-F")) %>%
  flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
  autofit() %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  colformat_num(j = 'Df', digits = 2) %>% 
  colformat_num(j = "Pseudo.F", digits = 2) %>%
  colformat_num(j = "p.value", digits = 4, na_str = "ns") %>% 
  align_nottext_col(align = "center", header = TRUE, footer = TRUE) %>% 
  align(align = "center", j = "p.value")

distDoc = read_docx()
distDoc = body_add_flextable(distDoc, value = distPermTab)
print(distDoc, target = "../tables/distTableVpi.docx")
distPermTab



```



<br><br>

# Plotting the data

***

To visualize the data and what our species composition looks like at our sites, we are going to plot the data in a few different ways. First we are going to just plot those with >2% abundance at our sites.

 Before we manipulate the data I am first making a vector called **myColors** using the function `gg_color_hue()` from the package `metafolio` that will save a vector of colors to the number of factors I have, this is so in our later plots, all the color palettes will be the same.


```{r, plotting, include = TRUE, eval = FALSE}

myColors <- brewer.pal(n = 26, name = "RdBu")

boxPlotData <- vpiMat1 %>% 
   rownames_to_column(var = "samples") %>% 
    separate(col = samples, into = c("date", "timePoint", "year", "site", "treatment", "reef"), sep = "_", remove = FALSE) %>%
  select(-c(samples)) %>% 
  gather(class, count, 7:32, factor_key = TRUE) %>% 
  as_tibble() %>% 
  mutate_at(c("date", "timePoint", "year", "site", "treatment", "reef", "class"), as.factor) %>% 
  group_by(timePoint, reef, year, treatment, class) %>% 
  summarise(sumCount = sum(count)) %>% 
  arrange(desc(sumCount)) %>% 
  group_by(timePoint, year, treatment) %>% 
  mutate(totalObservations = sum(sumCount)) %>% 
  group_by(timePoint, year, treatment, class) %>% 
  mutate(abundance = (sumCount/totalObservations)*100) %>% 
  # filter(abundance > 1) %>% 
  arrange(desc(abundance)) %>% 
  droplevels() %>% 
  mutate(class = factor(class, unique(class)))



vpiStack1 <- ggplot(boxPlotData, aes(x = treatment, y = abundance)) +
                      geom_bar(aes(fill = class), position = position_stack(), stat = 'identity', color = 'black', alpha = 0.8) +
                      ylim(0,100) +
                      labs(y = "Relative Abundance (%)", fill = "Class", title = "Relative Species Abundance > 1%")+
                      # scale_fill_manual(values = myColors)+
                      facet_wrap(timePoint ~ year, scales = 'free')




vpiStack <- vpiStack1+theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"), 
    # X Axis
    axis.text.x = element_text(size = 40, colour = "black", vjust = 0.5, hjust = 0.5, face= "bold"), 
    axis.title.x = element_blank(),
    # Y Axis
    axis.title.y = element_text(size = 40, face = "bold"), 
    axis.text.y = element_text(colour = "black", size = 40, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"), 
    axis.ticks = element_line(color="black"), 
    # Legend
    legend.title = element_text(size = 40, face = "bold"), 
    legend.text = element_text(size = 36, face = "bold", colour = "black"), 
    legend.position = "right",
    # Grid and Background
    panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = '#D3D3D3'), 
    # Text
    text = element_text(size=40, color="black"), 
    #rectangle
    rect = element_rect(fill = "transparent"))






vpiStack





ggsave("../figures/vpiStack.png", plot = vpiStack, width = 40, height = 20, units = 'in', dpi = 600)


```

# Coral Evenness, Richness, and coral cover

```{r, richnessAndEvenness, include = TRUE}
# Aggregate abundance by treatment and timePoint using dplyr
treatment_time_abundance <- vpiData %>%
  select(1:8) %>%
  filter(!class %in% c("Bare Substrate"))
 
# Calculate species richness
richness <- treatment_time_abundance %>%
  group_by(treatment, year, timePoint, reef, Site) %>%
  summarise(richness = specnumber(count))

richnessMat <- richness %>%
  tidyr::unite(sample, 1:4) %>% 
  spread(Site, richness, fill = 0) %>%
  column_to_rownames(var = 'sample')

samplesLookupPerm <- richnessMat %>% 
   rownames_to_column(var = "samples") %>% 
   select(samples) %>% 
   separate(col = samples, into = c("treatment", "year", "timePoint", "Reef"), sep = "_", remove = FALSE) %>% 
   column_to_rownames(var = "samples")


richnessDist <- vegdist(richnessMat, method = 'jaccard')

# Perform PERMANOVA
richnessPerm <- adonis2(richnessDist ~ treatment*timePoint*year, data = samplesLookupPerm, permutations = 9999)





# Calculate species evenness
evenness <- treatment_time_abundance %>%
  group_by(treatment, year, timePoint, reef, Site) %>%
  summarise(evenness = diversity(count, index = "simpson"))

evennessPerm <- adonis2(evenness$evenness ~ treatment * year * timePoint, data = evenness, permutations = 9999)





```
```{r, evenAndRichTables, include = TRUE}

richPermTab <- richnessPerm %>%
  as_tibble(rownames = "Comparison") %>%
  rename(`Pseudo.F` = `F`, `p.value` = `Pr(>F)`) %>% 
  select(Comparison, Df, `Pseudo.F`, `p.value`) %>% 
  mutate_if(is.numeric, round, digits = 3) %>%
    mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
  mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>% 
  mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–") %>%
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Df", value = as_paragraph("Df")) %>% 
  flextable::compose(part = "header", j = "Pseudo.F", value = as_paragraph("Pseudo-F")) %>%
  flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
  autofit() %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  colformat_num(j = 'Df', digits = 2) %>% 
  colformat_num(j = "Pseudo.F", digits = 2) %>%
  colformat_num(j = "p.value", digits = 4, na_str = "ns") %>% 
  align_nottext_col(align = "center", header = TRUE, footer = TRUE) %>% 
  align(align = "center", j = "p.value")

richDoc = read_docx()
richDoc = body_add_flextable(richDoc, value = richPermTab)
print(richDoc, target = "../tables/richnessTableVpi.docx")
richPermTab



evenPermTab <- evennessPerm %>%
  as_tibble(rownames = "Comparison") %>%
  rename(`Pseudo.F` = `F`, `p.value` = `Pr(>F)`) %>% 
  select(Comparison, Df, `Pseudo.F`, `p.value`) %>% 
  mutate_if(is.numeric, round, digits = 3) %>%
    mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
  mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>% 
  mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–") %>%
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Df", value = as_paragraph("Df")) %>% 
  flextable::compose(part = "header", j = "Pseudo.F", value = as_paragraph("Pseudo-F")) %>%
  flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
  autofit() %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  colformat_num(j = 'Df', digits = 2) %>% 
  colformat_num(j = "Pseudo.F", digits = 2) %>%
  colformat_num(j = "p.value", digits = 4, na_str = "ns") %>% 
  align_nottext_col(align = "center", header = TRUE, footer = TRUE) %>% 
  align(align = "center", j = "p.value")

evenDoc = read_docx()
evenDoc = body_add_flextable(evenDoc, value = evenPermTab)
print(evenDoc, target = "../tables/evennessTableVpi.docx")
evenPermTab








```







```{r, evenAndRichBoxPlots, include = TRUE}



richnessBox <- ggplot(richness, aes(x = timePoint, y = richness, fill = treatment)) +
  geom_bar(stat='identity', position = "dodge") +
  ylim(0, 10)+
  labs(title = "Species Richness by Treatment and Timepoint",
       x = "Timepoint",
       y = "Richness") +
  facet_wrap(treatment~site, scales = 'free_x')+
  theme_minimal()

evennessBox <- ggplot(evenness, aes(x = timePoint, y = evenness, fill = treatment)) +
  geom_boxplot() +
  labs(title = "Species Evenness by Treatment and Timepoint",
       x = "Timepoint",
       y = "Evenness") +
  theme_minimal()





```


```{r, coralCover, include = TRUE}
coralData <- vpiData %>% 
  filter(!class %in% c("Bare Substrate", "Chordata", "Crustose coralline algae" ,"Octocorallia", "Palythoa","Porifera", "Thalassia testudinum", "Turf algae","Xestospongia muta","Gorgonia ventalina", "Millepora spp.")) %>% 
  droplevels() 

coralMat <- coralData %>%
  select(-abundance) %>%
  tidyr::unite(sample, 1:6) %>% 
  spread(class, count, fill = 0) %>%
  column_to_rownames(var = 'sample')

samplesLookupPerm <- coralMat %>% 
   rownames_to_column(var = "samples") %>% 
   select(samples) %>% 
   separate(col = samples, into = c("date", "timePoint", "year", "site", "treatment", "reef"), sep = "_", remove = FALSE) %>% 
   column_to_rownames(var = "samples")



coralDist <- vegdist(coralMat, method = 'bray')

# Perform PERMANOVA
coralPerm <- adonis2(coralDist ~ treatment*timePoint*year, data = samplesLookupPerm, permutations = 9999)



```




```{r, coverGraph, include = TRUE}
coralStack1 <- ggplot(coralData, aes(x = timePoint, y = abundance)) +
                      geom_bar(aes(fill = class), position = position_stack(), stat = 'identity', color = 'black', alpha = 0.8) +
                      ylim(0,5) +
                      labs(y = "Relative Abundance (%)", fill = "Class", title = "Relative Coral Abundance")+
                      # scale_fill_manual(values = myColors)+
                      facet_wrap(treatment ~ year*Site, scales = 'free_x')

coralStack <- coralStack1+theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"), 
    # X Axis
    axis.text.x = element_text(size = 40, colour = "black", vjust = 0.5, hjust = 0.5, face= "bold"), 
    axis.title.x = element_blank(),
    # Y Axis
    axis.title.y = element_text(size = 40, face = "bold"), 
    axis.text.y = element_text(colour = "black", size = 40, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"), 
    axis.ticks = element_line(color="black"), 
    # Legend
    legend.title = element_text(size = 40, face = "bold"), 
    legend.text = element_text(size = 36, face = "bold", colour = "black"), 
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = '#D3D3D3'), 
    # Text
    text = element_text(size=40, color="black"), 
    #rectangle
    rect = element_rect(fill = "transparent"))
coralStack
ggsave("../figures/coverStack.png", plot = coralStack, width = 40, height = 20, units = 'in', dpi = 600)

```
```{r, coverPermTab, include = TRUE}




coralPermTab <- coralPerm %>%
  as_tibble(rownames = "Comparison") %>%
  rename(`Pseudo.F` = `F`, `p.value` = `Pr(>F)`) %>% 
  select(Comparison, Df, `Pseudo.F`, `p.value`) %>% 
  mutate_if(is.numeric, round, digits = 3) %>%
    mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
  mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>% 
  mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–") %>%
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Df", value = as_paragraph("Df")) %>% 
  flextable::compose(part = "header", j = "Pseudo.F", value = as_paragraph("Pseudo-F")) %>%
  flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
  autofit() %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  colformat_num(j = 'Df', digits = 2) %>% 
  colformat_num(j = "Pseudo.F", digits = 2) %>%
  colformat_num(j = "p.value", digits = 4, na_str = "ns") %>% 
  align_nottext_col(align = "center", header = TRUE, footer = TRUE) %>% 
  align(align = "center", j = "p.value")

coralDoc = read_docx()
coralDoc = body_add_flextable(coralDoc, value = coralPermTab)
print(coralDoc, target = "../tables/coralCoverTableVpi.docx")
coralPermTab



```











```{r, spongeAndAlgae, include = TRUE}

#Subsetting to assess changes in Macroalgae and Sponge cover


algae <- vpiData %>% 
  filter(class %in% c("Turf algae", "Xestospongia muta", "Porifera"), timePoint == c("T12")) %>% 
  droplevels()

kruskal.test(count ~ treatment, data = algae)











```



