---
title: "VPI"
author: "Ian Combs -- icombs@mote.org"
output:
  html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
library(magrittr)
```

### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

<!-- this is where the DOI would go  [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3675991.svg)](https://doi.org/10.5281/zenodo.3675991)
-->


#### [GitHub repository](https://github.com/icombs2017/photogrammetryNOAA.git){target="_blank"}

###
***
This is the analysis pipeline for data generated from TagLab annotations of our 6 10x10m plots.


***
### All analyses performe with R verion `r getRversion()`
# Basic setup of R environment
***
## Loading required packages

For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 

```{r, load packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2","officer","ggpubr", "rcompanion", "RColorBrewer", "patchwork", "magrittr","reshape2", "stringr", "plyr", "dplyr", "flextable", "tidyr", "tibble", "vegan", "forcats", "readr", "metafolio", "FSA","viridis", "paletteer")
pacman::p_load_gh("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
```

<br><br>

# Loading Data
***

Data is exported from Viscore's VPI scriplet as a .csv file and named according to the file. 
Since Viscore puts the "percent" column as a percentage sign (**%**) R cannot read it in  we are using `dplyr`'s `rename` function to change it to something R can read and that is easier to call and read when we are handling the data. Before we manipulate the data I am first making a vector called **myColors** using the function `gg_color_hue()` from the package `metafolio` that will save a vector of colors to the number of factors I have, this is so in our later plots, all the color palettes will be the same. I am also adding **site** and **time** columns to each subset, combining all subsets into one large dataframe using `bind_rows()`, adding a **treatment** column and using the `gsub()` to rename some 
```{r, load data, include = TRUE, results = 'hide'}



sitec1 <- as.factor("c1")
sitec2 <- as.factor("c2")
siteu1 <- as.factor("u1")
siteu2 <- as.factor("u2")
sitez1 <- as.factor("z1")
sitez2 <- as.factor("z2")
pre <- as.factor("Baseline")
post <- as.factor("1 Year")


vpiC1Pre <- read_csv("../data/vpi/2020_08_18_IC-C1pq.csv")
vpiC1Pre <- vpiC1Pre %>% 
  add_column(sitec1) %>% 
  rename(site = sitec1) %>%
  add_column(pre) %>% 
  rename(time = pre) %>% 
  dplyr::rename(abundance = `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))

vpiC2Pre <- read_csv("../data/vpi/2020_08_18_IC-C2pq.csv")
vpiC2Pre <- vpiC2Pre %>%
  add_column(sitec2) %>%
  rename(site = sitec2) %>% 
  add_column(pre) %>% 
  rename(time = pre) %>% 
  dplyr::rename(abundance = `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))


vpiU1Pre <- read_csv("../data/vpi/2020_08_19_IC-U1pq.csv")
vpiU1Pre <- vpiU1Pre %>%
  add_column(siteu1) %>%
  rename(site = siteu1) %>%
  add_column(pre) %>% 
  rename(time = pre) %>% 
  dplyr::rename(abundance =  `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))

vpiU2Pre <- read_csv("../data/vpi/2020_08_19_IC-U2pq.csv")
vpiU2Pre <- vpiU2Pre %>%
  add_column(siteu2) %>% 
  rename(site = siteu2) %>%
  add_column(pre) %>% 
  rename(time = pre) %>% 
  dplyr::rename(abundance =  `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))

vpiZ1Pre <- read_csv("../data/vpi/2020_08_18_IC-Z1pq.csv")
vpiZ1Pre <- vpiZ1Pre %>%
  add_column(sitez1) %>%
  rename(site = sitez1) %>%
  add_column(pre) %>% 
  rename(time = pre) %>% 
  dplyr::rename(abundance =  `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))

vpiZ2Pre <- read_csv("../data/vpi/2020_08_18_IC-Z2pq.csv")
vpiZ2Pre <- vpiZ2Pre %>% 
  add_column(sitez2) %>% 
  rename(site = sitez2) %>%
  add_column(pre) %>% 
  rename(time = pre) %>% 
  dplyr::rename(abundance =  `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))

vpiC1Post <- read_csv("../data/vpi/2021_12_08_IC-C1pq.csv")
vpiC1Post <- vpiC1Post %>%
  add_column(sitec1) %>%
  rename(site = sitec1) %>%
  add_column(post) %>% 
  rename(time = post) %>% 
  dplyr::rename(abundance = `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))

vpiC2Post <- read_csv("../data/vpi/2021_12_07_IC-C2pq.csv")
vpiC2Post <- vpiC2Post %>% 
  add_column(sitec2) %>% 
  rename(site = sitec2) %>%
  add_column(post) %>% 
  rename(time = post) %>%
  dplyr::rename(abundance = `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))

vpiU1Post <- read_csv("../data/vpi/2021_12_08_IC-U1pq.csv")
vpiU1Post <- vpiU1Post %>%
  add_column(siteu1) %>%
  rename(site = siteu1) %>%
  add_column(post) %>% 
  rename(time = post) %>%
  dplyr::rename(abundance =  `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))

vpiU2Post <- read_csv("../data/vpi/2021_12_08_IC-U2pq.csv")
vpiU2Post <- vpiU2Post %>%
  add_column(siteu2) %>% 
  rename(site = siteu2) %>%
  add_column(post) %>% 
  rename(time = post) %>%
  dplyr::rename(abundance =  `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))

vpiZ1Post <- read_csv("../data/vpi/2021_12_07_IC-Z1pq.csv")
vpiZ1Post <- vpiZ1Post %>%
  add_column(sitez1) %>%
  rename(site = sitez1) %>%
  add_column(post) %>% 
  rename(time = post) %>%
  dplyr::rename(abundance =  `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))

vpiZ2Post <- read_csv("../data/vpi/2021_12_07_IC-Z2pq.csv")
vpiZ2Post <- vpiZ2Post %>%
  add_column(sitez2) %>%
  rename(site = sitez2) %>%
  add_column(post) %>% 
  rename(time = post) %>%
  dplyr::rename(abundance =  `%`) %>% 
  na.omit() %>% 
  arrange(desc(abundance))

vpi <- bind_rows(vpiC1Pre, vpiC1Post, vpiC2Pre, vpiC2Post, vpiU1Pre, vpiU1Post, vpiU2Pre, vpiU2Post, vpiZ1Pre, vpiZ1Post, vpiZ2Pre, vpiZ2Post)
vpi <- as_tibble(vpi)


vpi$class <- gsub("Crustose coralline algae natans","Crustose coralline algae",vpi$class)
vpi$class <- gsub("Rubble", "Bare Substrate", vpi$class)
vpi$class <- as.factor(vpi$class)

vpi$treatment = if_else(vpi$site %in% c("c2", "z1", "u1"), "Control","Outplanted") 
vpi$reef = if_else(vpi$site %in% c("c1", "c2"), "Site C",
                   if_else(vpi$site %in% c("u1", "u2"), "Site U", "Site Z"))
vpi$reef <- as.factor(vpi$reef)

myColors <- gg_color_hue(length(vpi$class))






```

<br><br>
# Plotting the data
***

To visualize the data and what our species composition looks like at our sites, we are going to plot the data in a few different ways. First we are going to just plot those with >2% abundance at our sites.




```{r}
myColors <- paletteer_c("harrypotter::gryffindor", 6, direction = 1 )
#myColors <- paletteer_c("pals::ocean.balance", 6, direction = 1)


# Using this just to plot the top 2% of species
vpi1 <- vpi %>% 
  filter(abundance > 2) %>% 
  droplevels()
vpi1$class = factor(vpi1$class, levels(vpi1$class)[c(2,6,3,4,5,1)])



vpiStack1 <- ggplot(vpi1, aes(x = site, y = abundance, fill = class, )) + 
    geom_bar(stat = "identity", colour = "black", position = position_stack(reverse = TRUE))+
    ylim(0,100) + 
    labs(y = "Relative Abundance (%)", fill = "Class", title = "Relative Species Abundance > 2%")+
    scale_fill_manual(values = myColors)+
    facet_wrap(treatment~ time , scales = 'free_x')




vpiStack

ggsave("../figures/noaaReportFig2.jpeg", plot = vpiPreStack, width = 20, height = 15, units = "in", dpi = 600)





vpiStack1 <- ggplot(vpi1, aes(x = reef, y = abundance, fill = class, group = reef))+
                      geom_bar(stat = 'identity', color = 'black', position = position_stack(reverse = TRUE))+
                      ylim(0,100)+
                      labs(y = "Relative Abundance (%)", fill = "Class", title = "Relative Species Abundance > 2%")+
                      scale_fill_manual(values = myColors)+
                      facet_grid(treatment ~ time)
vpiStack <- vpiStack1+theme(
    axis.text.x = element_text(size = 40, colour = "black", vjust = 0.5, hjust = 0.5, face= "bold"), 
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 40, face = "bold"), 
    axis.text.y = element_text(colour = "black", size = 40, face = "bold"),
    legend.title = element_text(size = 40, face = "bold"), 
    legend.text = element_text(size = 36, face = "bold", colour = "black"), 
    panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
    panel.background = element_rect(fill = '#F5F5F5'), 
    plot.title = element_text(size = 40, face = "bold"), 
    axis.line = element_line(colour = "black"), 
    axis.ticks = element_line(color="black"), 
    text = element_text(size=40, color="black"), 
    legend.position = "bottom")
vpiStack

ggsave("../figures/noaaReportFig2.jpeg", plot = vpiPreStack, width = 20, height = 15, units = "in", dpi = 600)





```


<br><br>

# Normality Test
***

Conducting a few normality tests to see whether or not the data violates the assumptions of normality. And then running the appropriate statistical test. In this case we are running a single-factor Kruskal-Wallis (for non-normally distributed data) and then a multi-factor Permuational Analysis of Variance (PERMANOVA) using the function `adonis` in the package `vegan` looking at abundance across treatment (control vs outplanted) and functional groups (gorgonians, corals, etc). The factors and variables are likely to change as we add in variables from other data sets (potentially) and timepoints (definitely). 

```{r}

vpiPerm <- adonis(formula = abundance ~ time*treatment*class, data = vpi, method = "euclidian", permutations = 9999)

vpiPairwise <- pairwise.adonis(vpi[c(3)], factors = vpi$class, sim.method = 'euclidian', p.adjust.m = 'bonferroni', perm = 9999)

vpiPairwise <- vpiPairwise %>% filter(p.adjusted < 0.05)

wilcox.test(count ~ class, data = vpi)
vpiDunn <- dunnTest(count ~ class, data = vpi, method = 'bh')
vpiDunn <- vpiDunn$res %>% filter(P.adj < 0.05)






```












```{r, plotting, include = TRUE}


vpiC1PrePlot1 <- ggplot(vpiC1Pre, aes(x = fct_reorder(class, -abundance), y = abundance, fill = class))+
              geom_bar(stat = 'identity')+
              labs(y = "Percent Abundance", x = "C1",title = "Pre-Restoration")+
              ylim(0,100)+
              scale_color_brewer(palette = "Set3")
vpiC1PrePlot <- vpiC1PrePlot1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'), 
  plot.title = element_text(size = 15, face = "bold"), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size=20, color="black"), 
  axis.text.x=element_text(angle = 45, hjust = 1, size=14, color="black"), 
  axis.text.y=element_text(size=12, color="black"),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank()
)+rremove("legend")

vpiC1PrePlot

vpiC2PrePlot1 <- ggplot(vpiC2Pre, aes(x = fct_reorder(class, -abundance), y = abundance, fill = class))+
              geom_bar(stat = 'identity')+
              labs(y = "Percent Abundance", x = "C2",title = "Pre-Restoration")+
              ylim(0,100)+
              scale_color_brewer(palette = "Set3")
vpiC2PrePlot <- vpiC2PrePlot1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'), 
  plot.title = element_text(size = 15, face = "bold"), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size=20, color="black"), 
  axis.text.x=element_text(angle = 45, hjust = 1, size=14, color="black"), 
  axis.text.y=element_text(size=12, color="black"),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank()
)+rremove("legend")

vpiC2PrePlot

vpiU1PrePlot1 <- ggplot(vpiU1Pre, aes(x = fct_reorder(class, -abundance), y = abundance, fill = class))+
              geom_bar(stat = 'identity')+
              labs(y = "Percent Abundance", x = "U1",title = "Pre-Restoration")+
              ylim(0,100)+
              scale_color_brewer(palette = "Set3")
vpiU1PrePlot <- vpiU1PrePlot1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'), 
  plot.title = element_text(size = 15, face = "bold"), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size=20, color="black"), 
  axis.text.x=element_text(angle = 45, hjust = 1, size=14, color="black"), 
  axis.text.y=element_text(size=12, color="black"),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank()
)+rremove("legend")

vpiU1PrePlot

vpiU2PrePlot1 <- ggplot(vpiU2Pre, aes(x = fct_reorder(class, -abundance), y = abundance, fill = class))+
              geom_bar(stat = 'identity')+
              labs(y = "Percent Abundance", x = "U2",title = "Pre-Restoration")+
              ylim(0,100)+
              scale_color_brewer(palette = "Set3")
vpiU2PrePlot <- vpiU2PrePlot1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'), 
  plot.title = element_text(size = 15, face = "bold"), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size=20, color="black"), 
  axis.text.x=element_text(angle = 45, hjust = 1, size=14, color="black"), 
  axis.text.y=element_text(size=12, color="black"),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank()
)+rremove("legend")

vpiU2PrePlot


vpiZ1PrePlot1 <- ggplot(vpiZ1Pre, aes(x = fct_reorder(class, -abundance), y = abundance, fill = class))+
              geom_bar(stat = 'identity')+
              labs(y = "Percent Abundance", x = "Z1",title = "Pre-Restoration")+
              ylim(0,100)+
              scale_color_brewer(palette = "Set3")
vpiZ1PrePlot <- vpiZ1PrePlot1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'), 
  plot.title = element_text(size = 15, face = "bold"), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size=20, color="black"), 
  axis.text.x=element_text(angle = 45, hjust = 1, size=14, color="black"), 
  axis.text.y=element_text(size=12, color="black"),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank()
)+rremove("legend")

vpiZ1PrePlot



vpiZ2PrePlot1 <- ggplot(vpiZ2Pre, aes(x = fct_reorder(class, -abundance), y = abundance, fill = class))+
              geom_bar(stat = 'identity')+
              labs(y = "Percent Abundance", x = "Z2",title = "Pre-Restoration")+
              ylim(0,100)+
              scale_color_brewer(palette = "Set3")
vpiZ2PrePlot <- vpiZ2PrePlot1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'), 
  plot.title = element_text(size = 15, face = "bold"), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size=20, color="black"), 
  axis.text.x=element_text(angle = 45, hjust = 1, size=14, color="black"), 
  axis.text.y=element_text(size=12, color="black"),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank()
)+rremove("legend")

vpiZ2PrePlot


compPlotPre <- vpiC1PrePlot + vpiC2PrePlot + vpiU1PrePlot + vpiU2PrePlot + vpiZ1PrePlot + vpiZ2PrePlot + plot_annotation(
  title = 'Baseline Benthic Community Composition using Virtual Point Intercept',
  theme = theme(plot.title = element_text(size = 18)))


ggsave("../figures/vpiCompPlotPre.png", plot = compPlotPre, width = 15, height = 10, units = "in", dpi = 300)






```





























```{r, plotting, include = TRUE}


vpiPrePlot1 <- ggplot(vpiC1Pre, aes(x = fct_reorder(class, -abundance), y = abundance, fill = class))+
              geom_bar(stat = 'identity')+
              labs(y = "Percent Abundance", x = "",title = "Pre-Restoration")+
              ylim(0,60)
vpiPrePlot <- vpiPrePlot1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'), 
  plot.title = element_text(size = 15, face = "bold"), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size=20, color="black"), 
  axis.text.x=element_text(angle = 45, hjust = 1, size=14, color="black"), 
  axis.text.y=element_text(size=12, color="black"),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank()
)+rremove("legend")

vpiPrePlot

vpiPostPlot1 <- ggplot(vpiC1Post, aes(x = fct_reorder(class, -abundance), y = abundance, fill = class))+
              geom_bar(stat = 'identity')+
              labs(y = "Percent Abundance", x = "Class", title = "Post-Restoration")+
              ylim(0,60)


vpiPostPlot <- vpiPostPlot1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'), 
  plot.title = element_text(size = 15, face = "bold"), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size=20, color="black"), 
  axis.text.x=element_text(angle = 45, hjust = 1, size=14, color="black"), 
  axis.text.y=element_text(size=12, color="black"),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank()
)+rremove("legend")

vpiPostPlot

compPlot <- vpiPrePlot / vpiPostPlot + plot_annotation(
  title = 'Benthic Community Composition Pre- and Post- Restoration using Virtual Point Intercept',
  theme = theme(plot.title = element_text(size = 18)))
ggsave("../figures/vpiCompPlot.png", plot = compPlot, width = 15, height = 10, units = "in", dpi = 300)
```

